---
title: "Time Aggregation"
author: "Daniel Klotz"
---


```{r setup10, include=FALSE, purl=FALSE}
  knitr::opts_chunk$set(eval = FALSE, tidy = FALSE, fig.align = 'center')
```

In hydrology it is often useful to summarize the data respect to a
given time dimension. In visCOS this can be done by using the
`aggregate_time` function. The function takes in [cosdata](cook_data.html) and
aggregates them according to a chosen time dimension. In the following are
some examples of its use:

```{r, eval=TRUE, purl=FALSE}
library(coscos)
library(magrittr)
library(visCOS)
library(ggplot2)
library(dplyr)
library(pasta)
```

Daily runoff aggregation:
```{r, eval=TRUE, purl=FALSE}
  cosdata <- coscos::viscos_example()
  runoff_aggregate_dd <- aggregate_time(cosdata, "dd") %>%
    dplyr::mutate(obs_sim = substr(key,1,4),
                  basin = substr(key,6,9) %>% as.integer(.))
  # plot data:
  ggplot(runoff_aggregate_dd) +
      geom_line(aes(x = dd, y = value, col = obs_sim)) +
      scale_colour_manual(values = c(viscos_options("color_o"),
                                     viscos_options("color_s"))) +
      facet_wrap( ~ basin,ncol = 1) +
      theme_minimal()
```

Monthly runoff aggregation:
```{r, eval = TRUE, purl=FALSE}
  runoff_aggregate_mm <- aggregate_time(cosdata, "mm") %>%
    dplyr::mutate(obs_sim = substr(key,1,4),
                  basin = substr(key,6,9) %>% as.integer(.))
  # plot data:
  ggplot(runoff_aggregate_mm) +
      geom_line(aes(x = mm, y = value, col = obs_sim)) +
      scale_colour_manual(values = c(viscos_options("color_o"),
                                     viscos_options("color_s"))) +
      scale_x_discrete(limits = runoff_aggregate_mm$mm) +
      facet_wrap( ~ basin) +
      theme_minimal()
```

Yearly runoff aggregation:
```{r, eval = TRUE, purl=FALSE}
  runoff_aggregate_yyyy <- aggregate_time(cosdata, "yyyy") %>%
    dplyr::mutate(obs_sim = substr(key,1,4),
                  basin = substr(key,6,9) %>% as.integer(.))
  # plot data:
  ggplot(runoff_aggregate_yyyy) +
    geom_line(aes(x = yyyy, y = value, col = obs_sim)) +
    geom_point(aes(x = yyyy, y = value, col = obs_sim)) +
    scale_colour_manual(values = c(viscos_options("color_o"),
                                   viscos_options("color_s"))) +
    facet_wrap( ~ basin) +
    scale_x_discrete(limits = runoff_aggregate_yyyy$yyyy,  
                    labels = abbreviate) +
    theme_minimal()
```

Yearly and monthly runoff aggregation:
```{r, eval = TRUE, purl=FALSE}
  runoff_aggregate_yyyymm <- aggregate_time(cosdata, c("yyyy","mm")) %>%
    dplyr::mutate(obs_sim = substr(key,1,4),
                  basin = substr(key,6,9) %>% as.integer(.),
                  idx = yyyy + mm/12) # a small hack to get a unique identifier
  # plot data:
  ggplot(runoff_aggregate_yyyymm) +
    geom_line(aes(x = idx, y = value, col = obs_sim)) +
    scale_colour_manual(values = c(viscos_options("color_o"),
                                   viscos_options("color_s"))) +
    scale_x_discrete(limits = runoff_aggregate_yyyymm$idx,  
                     labels = runoff_aggregate_yyyymm$posixdate) +
    facet_wrap( ~ basin,ncol = 1) +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 90, hjust = 1))
```
