---
title: "Runoff Peaks"
author: "Daniel Klotz"
---

```{r setup8, include=FALSE, purl=FALSE}
  knitr::opts_chunk$set(eval = FALSE, tidy = FALSE, fig.align = 'center')
```

The function `peack` lets users explore the highest events in
among the available basins. It provides a list of [ggplot2](http://ggplot2.org/)
plots, containing an overview plot (`overview`), a scatter plot (`scatter`)
and detail plots of the individual events (`event_plot`). Instead of
explaining the properties of each plot in detail it is best to get an intuition
of the function by looking at some examples.


## Examples
For the examples 10 events are extracted from a runoff example
```{r, eval = TRUE, purl=FALSE}
  library(coscos)
  library(visCOS)
  cosdata <- viscos_example()
  peakplots <- peack(cosdata, action = "plots", n_events = 10L)

```

The `peakplots` list does now contain plots for each basin within the
`cosdata` data.frame:
```{r, eval = TRUE, purl=FALSE}
  names(peakplots)
```

For each basin the a set of plots (`overview`,`scatter`,`event_plot`) are saved
within a list for each basin. In the following the plots for basin 1 are shown:
```{r, eval = TRUE, purl=FALSE}
  names(peakplots$basin0001)
```

The `overview` plot shows the entire time series of `data1` and `data2` of the
basin. The found events are marked with black dots. The `overview` plot for
basin 1 is:
```{r, eval = TRUE, purl=FALSE}
  peakplots$basin0001$overview
```

The `scatter` plot shows the found events within a scatter plot, where  `data1`
is the x-axis and `data2` on the y-axis. In the following an example for basin 1
is given.
```{r, eval = TRUE, purl=FALSE}
  peakplots$basin0001$scatter
```

Detail plots for each of the found events are given in form of the `event_plot`
objects. Here an example:
```{r, eval = TRUE, purl=FALSE}
  peakplots$basin0001$event_plot5
```


## References
- http://stats.stackexchange.com/questions/22974/how-to-find-local-peaks-valleys-in-a-series-of-data (checked 12/2016)
- http://stats.stackexchange.com/questions/36309/how-do-i-find-peaks-in-a-dataset (checked 12/2016)
