---
title: "Judging with Objective Functions"
author: "Daniel Klotz"
---


```{r setup6, include=FALSE, purl=FALSE}
  knitr::opts_chunk$set(eval = FALSE, tidy = FALSE)
```

This chapter presents a set of functions to organize and/or visualize a
set of objective functions that ought to be evaluated over all basins. The
objective functions can be provided as a list via `.ofuns`. The pre-set
objective functions are (these are provided by the `coscos` package, see also:
[distance metrics](LP-distance_measures.html)):

- Nash-Sutcliffe Efficency: `of_nse`
- Kling-Gupta Efficency: `of_kge`
- Percentage Bias: `of_pbias`
- Pearson's Correlation Coefficient: `of_corr`

The function `judge_compute` applies the provided objective functions to all
basins. The computation is done for the overall time-span and for the
*marked periods*. On top of that two functions for summary plots are implemented:
`judge_rasterplot` and `judge_barplot`. A notable quirk of both functions is
that they return a list as output (whereby each entry or sub-entry of the list
contains a `ggplot2` figure). Two interactive tools for exploration are provided
by the `judge_explore` and `judge_explore2` functions. Finally, the `judge`
function wraps for all other `judge_` functions. The selection is done via the
`type` argument (the variable is evaluated in a non-standard fashion, which
implies, among other things, that the argument can be provided without it
being a character). Here are some examples:

```{r, message=3, purl=FALSE, eval=TRUE}
library(magrittr)
library(coscos)
library(visCOS)
library(pasta)
```

**Example 1.** Evaluating the `judge` function without arguments executes
`judge_compute`, which returns a data.frame with the objective functions for the
entire time-series and the marked periods (of each basin):
```{r, message=3, purl=FALSE, eval=TRUE}
viscos_example( ) %>%
  judge(.) %>%
  print(.)
```

**Example 2.** Plotting the computed objective functions as bar-plots:
```{r, message=3, purl=FALSE, eval=TRUE}
viscos_example() %>%
  judge(., barplot) %>%
  .[[1]] %>% # a list with the plots is returned; select first entry
  plot(.)
```

**Example 3.** Using the `judge_` functionality:
```{r, message=3, purl=FALSE, eval=TRUE}
viscos_example() %>%
  judge_barplot(.) %>%
  .[[1]] %>% # a list with the plots is returned; select one entry
  plot(.)
```

**Example 4.** Plotting the results of main objective function with a raster:
```{r, message=3, purl=FALSE, eval=FALSE}
viscos_example() %>%
  judge(., "rasterplot") %>%
  .[[2]] %>%
  plot(.)
```
