---
title: "Flow Duration Curves"
author: "Daniel Klotz"
---


```{r setup4, include=FALSE, purl=FALSE}
  knitr::opts_chunk$set(eval = FALSE, tidy = FALSE)
```

Flow-duration curves represent the relationship between the magnitude and the
frequency of a streamflow. They provide an estimate of the percentage of time
a given streamflow was exceeded within the evaluated time frame. Foster [1934]
attributes the first use of flow duration curves to Clemens Herschel around 1880.
Since they have been used for a wide array of applications.
visCOS provides a function to compute the data for flow-duration curves and
a function to plot them directly.
The former function is called `fdc_compute`. It computes the flow exceedance
properties and returns a data.frame. The calculations are adapted from the
method used within the [hydroTSM](https://CRAN.R-project.org/package=hydroTSM)
package. It is currently rather slow.
The plot-function is called `fdc_plot`. Internally it uses `fdc_compute` for the
data preparation and generates a facetted `ggplot` object from it. In the plot
each basin is a facet and each sub-plot shows the $o$-data and $s$-data (see:
[Introduction](Introduction.html)).


## Example
Flow duration curves can be plotted in visCOSin the following way:
```{r example, purl=FALSE, eval=TRUE}
  library(coscos)
  library(visCOS)
  cosdata <- coscos::viscos_example()
  fdc_plot(cosdata)
```


## References
- Foster, H.A. (1934): Duration curves. ASCE Trans., 99, 1213-1267
- Vogel, R. M.; Fennessy, N. M. (1994): Flow-Duration Curves. I: New Interpretation
and Confidence Intervals. JWRMD 120, No. 4
