# cosdata
For visCOS raw-data is data directly taken from COSERO. In general, raw-data
can have many forms. The focus of visCOS are time series of observations, $o$,
and model simulations, $s$. It assumes that the data is already loaded in to R.

R has many options to read data. Most file-formats can be accessed with the
native `read.` functions (for an example enter `?read.table` in your R-terminal).
However, for larger files we recommend the read functionality of the
`data.table` package. In our tests it was fastest and most flexible choice.
Lastly, we recommend `readr` package for well structured and large data.

After the data is read into R it needs to be transformed (*cooked*) into a
rigidly defined format, which we call **cosdata**. The format has some redundant
columns, but provides a large optionality. It is also a good middle ground
between the output of COSERO and the formats used within the R ecosystem.

All cosdata formated data can be used for further exploratory analysis with
visCOS.

**NOW WE NEED TO EXPLAIN COSDATA!**

Snipptes:


Currently visCOSonly allows to compare between *numbered catchments*! The
data **must** include an integer number at the end of its name, e.g.
`QObs_001` and `QSim_001`).

Additionally, `cosdata` data.frame needs to have two different definitions for
the date of a given row (see: [Introduction](introduction.html)). One is based
on an old way to write dates out in fortran code. In that case each column
represents a time-resolution. In concrete, the following columns are used:

-  *yyyy* - year,
- *mm* - month,
- *hh* - hour,
- *min* - minute.

The other format is a more modern way to define time information. That is,
the `POSIXct` format (see: [link](https://stat.ethz.ch/R-manual/R-devel/library/base/html/as.POSIXlt.html)).
This format is a standard for `R` and has many usages (e.g. transforming your
data frame into a time series).

visCOS can differentiate seasonal information on a monthly
resolution. The in- and out-of-period markings are stored in a separate column
(defined by `viscos_options("name_COSperiod")`). Here, each season is defined
by an number (integer), which starts 1 and is raised for each new season. A 0
indicates the out-of-period rows.

If needed, the column-names of the cosdata can adapted by using the
`viscos_options` function from `coscos` (alternatively one might prefer to set
the column-name the `data.frame` accordingly). Calling the function
`viscos_options()` will return a list with the available options. The available
options are as following:
```{r, purl=FALSE, eval=TRUE}
   # showing the options:
   library(coscos)
   library(magrittr)
   viscos_options() %>% unlist(.)
```

An option can be changed by providing the new option-value as functional
argument of the option. Here is an example for changing the $o$ column color:
```{r, purl=FALSE, eval=TRUE}
   viscos_options(color_o = "dodgerblue")
   viscos_options( ) %>% unlist(.)
```

The `coscos` package also provides a function to get some exemplary data to use
in visCOS. Throughout this guide we will use it to demonstrate the different
functionalities of visCOS. The data can be loaded with the function
`viscos_example`, e.g.:
```{r, purl=FALSE, eval=TRUE, message=3:5}
  options(width=80)
  require(coscos)
  #
  raw_data_example <- viscos_example( )
  head(raw_data_example)
```

One can see that the example data still has some columns that are not needed for
further analysis, i.e. the columns `QOSI_0001` and `QOSI_0002`. Furthermore,
the `posixdate` and the `period` columns are still missing in the data. Again,
the `coscos` provides a streamlined function, called `cook_cosdata` to get rid
of the non-needed columns, complete the date columns and add the `period`
column with pre-specified settings. With the raw-data example from before, the
finished cosdata might then look as following:
```{r, purl=FALSE, eval=TRUE, message=3:5}
  raw_data_example %>%
    cook_cosdata(.) %>%
    head(.)
```

The `cook_cosdata` function can be very useful. If you are not sure that your
data has the right format yet, just try to use `cook_cosdata` on it. The function
is used at the start of almost all visCOS functions. Thus, it is very likely
that your data is fine if `cook_cosdata` works on it.

## Examples plots from cosdata
It is already possible to do a lot with just the cosdata. In the following are
some examples:
```{r, purl=FALSE, eval=TRUE}
  library(ggplot2)
  raw_data_example %>%
    cook_cosdata(.) %>%
    ggplot(.) +
      geom_line(aes(x = posixdate, y = QOBS_0002),
                color = viscos_options("color_o")) +
       geom_line(aes(x = posixdate, y = QSIM_0002),
                color = viscos_options("color_s"))   
```

```{r, purl=FALSE, eval=TRUE}
  library(ggplot2)
  raw_data_example %>%
    cook_cosdata(.) %>%
    ggplot(.) +
      geom_point(aes(x = QOBS_0001, y = QOBS_0002))  
```
