# Objective Functions

```{r setup6, include=FALSE, purl=FALSE}
  knitr::opts_chunk$set(eval = FALSE, tidy = FALSE)
```

This chapter presents a set of functions to organize and/or visualize a
set of objective functions that ought to be evaluated over all basins. The
objective functions can be provided as a list via `.ofuns`. The pre-set
objective functions are (these are provided by the `coscos` package, see also:
[distance metrics](LP-distance_measures.html)):

- Nash-Sutcliffe Efficency: `of_nse`
- Kling-Gupta Efficency: `of_kge`
- Percentage Bias: `of_pbias`
- Pearson's Correlation Coefficient: `of_corr`

The function `judge_compute` applies the provided objective functions to all
basins. The computation is done for the overall time-span and for the
*marked periods*. On top of that two functions for summary plots are implemented:
`judge_rasterplot` and `judge_barplot`. A notable quirk of both functions is
that they return a list as output (whereby each entry or sub-entry of the list
contains a `ggplot2` figure). Two interactive tools for exploration are provided
by the `judge_explore` and `judge_explore2` functions. Finally, the `judge`
function wraps for all other `judge_` functions. The selection is done via the
`type` argument (the variable is evaluated in a non-standard fashion, which
implies, among other things, that the argument can be provided without it
being a character). Here are some examples:

```{r, message=3, purl=FALSE, eval=TRUE}
library(magrittr)
library(coscos)
library(visCOS)
```

**Example 1.** Evaluating the `judge` function without arguments executes
`judge_compute`, which returns a data.frame with the objective functions for the
entire time-series and the marked periods (of each basin):
```{r, message=3, purl=FALSE, eval=TRUE}
viscos_example( ) %>%
  judge(.) %>%
  print(.)
```

**Example 2.** Plotting the computed objective functions as bar-plots:
```{r, message=3, purl=FALSE, eval=TRUE}
viscos_example() %>%
  judge(., barplot) %>%
  .[[1]] %>% # a list with the plots is returned; select first entry
  plot(.)
```

**Example 3.** Using the `judge_` functionality:
```{r, message=3, purl=FALSE, eval=TRUE}
viscos_example() %>%
  judge_barplot(.) %>%
  .[[1]] %>% # a list with the plots is returned; select one entry
  plot(.)
```

**Example 4.** Plotting the results of main objective function with a raster:
```{r, message=3, purl=FALSE, eval=TRUE}
viscos_example() %>%
  judge(., "rasterplot") %>%
  .[[2]] %>%
  plot(.)
```


## Code
```{r}
# ---------------------------------------------------------------------------
# Code for the the judge functions
# authors: Daniel Klotz, Johannes Wesemann, Mathew Herrnegger
# !!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!

```

### Wrapper
```{r}
#' Wrapper for the different judge functions
#'
#' A wrapper for all `judge_` functions. The
#' input arguments `type` is lazily-evaluated (which basically implies that
#' it does not have to be provided as a character argument).
#' which function is chosen (the "pre fix" part of
#' the selected functions, i.e. the `judge_` prefix does not need to be provided).
#'
#' @import pasta
#' @import lazyeval
#'
#' @rdname judge
#' @export
judge <- function(cosdata,
                  type = "compute",
                  .ofuns = list(nse   = coscos::of_nse,
                                kge   = coscos::of_kge,
                                pbias = coscos::of_pbias,
                                corr  = coscos::of_cor),
                  opts = coscos::viscos_options(),
                  ...
                  ) {
  # lazy evaluation: ------------------------------------------------------
  le_dots <-  match.call(expand.dots = FALSE) %>%
    .["..."] %>%
    .[[1]]
  le_type <- match.call(expand.dots = FALSE) %>%
    .["type"] %>%
    .[[1]] %>%
    as.character(.)
  if (length(le_type) == 0) {
    le_type = "compute"
  }
  if (class(le_type) != "character") {
    stop("Cannot resolve 'type' argument. Try to provide a character.")
  }
  # compute options: ------------------------------------------------------
  if (le_type == "compute") {
    judge_compute(cosdata, .ofuns)

  } else if (le_type == "judge_barplot" | le_type == "barplot") {
    judge_barplot(cosdata, .ofuns)

  } else if (le_type == "judge_rasterplot" | le_type == "rasterplot") {
    judge_rasterplot(cosdata, .ofuns)

  } else if (le_type == "judge_explore" | le_type == "explore") {
    judge_explore(cosdata, .ofuns)

  } else if (le_type == "judge_explore" | le_type == "compare") {
    cosdata2 <- NULL
    if ("cosdata2" %in% names(le_dots)) {2 <- eval(le_dots$cosdata2)}
    judge_compare(cosdata, cosdata2, .ofuns)

  } else if (le_type == "judge_window" | le_type == "window") {
    windows <- c(20L,500L)
    if (window_sizes %in% names(le_dots)) {windows <- window_sizes}
    judge_window(cosdata, .ofuns, window_sizes)
  } else {
    stop("There is no option (`type`) called" %&&% le_type)
  }
}
```



### Computation
The purpose of this function is to compute a set of objective functions for `cosdata`.

```{r}
# ---------------------------------------------------------------------------
#' Get basic objective function for
#'
#' Calculate basic objective functions(NSE, KGE, percentage BIAS, Correlation)
#' for every basin and the chosen periods.
#'
#' @param cosdata cosdata data.frame.
#' @return list of basic objective function evaluated for the different
#' hydrological years and over the whole timespan.
#'
#' @import pasta
#' @import tibble
#' @importFrom purrr map_df
#' @importFrom dplyr select filter starts_with
#' @importFrom magrittr set_names
#' @importFrom tibble as_tibble
judge_compute <- function(cosdata,
                          .ofuns = list(nse   = coscos::of_nse,
                                        kge   = coscos::of_kge,
                                        pbias = coscos::of_pbias,
                                        corr  = coscos::of_cor),
                          opts = coscos::viscos_options()
                          ) {
  # pre: ====================================================================
  cos_data <- coscos::cook_cosdata(cosdata)
  name_period <- opts[["name_COSperiod"]]
  name_o <- opts[["name_o"]]
  name_s <- opts[["name_s"]]
  #
  if( !(class(.ofuns) == "list") ) {
    .ofuns <- list(.ofuns)
  }
  if (is.null(names(.ofuns))){
    d_names <- "d" %&% 1:length(.ofuns)
  } else {
    d_names <- names(.ofuns)
  }
  #
  evaluation_data <- cos_data[cos_data[[name_period]] > 0, ]
  number_judge_basins <- evaluation_data %>%
    names(.) %>%
    unique(.) %>%
    grepl(name_o, ., ignore.case = TRUE) %>%
    sum(.)
  data_periods <- evaluation_data %>%
    .[[name_period]] %>%
    unique(.)
  number_judge_periods <- data_periods %>% length

  # compute main-of for entire data: ========================================
  d_mean <- purrr::map_df(.ofuns, function(judge_, x, y) as.numeric(judge_(x,y)),
                 x = dplyr::select(evaluation_data,dplyr::starts_with(name_o)) %>% unname(.),
                 y = dplyr::select(evaluation_data,dplyr::starts_with(name_s)) %>% unname(.) ) %>%
    t(.) %>%
    as_tibble(.) %>%
    magrittr::set_names(., "basin" %_% 1:number_judge_basins) %>%
    cbind.data.frame(of = d_names,.)
  # compute periodwise main-of: =============================================
  period_compute <- function(k) {
    o_pick <- dplyr::filter(evaluation_data,period == data_periods[k]) %>%
      dplyr::select(.,starts_with(name_o)) %>%
      unname(.)
    s_pick <- dplyr::filter(evaluation_data,period == data_periods[k]) %>%
      dplyr::select(.,starts_with(name_s)) %>%
      unname(.)
    d_measures <- purrr::map_df(.ofuns, function(judge_,x,y) as.numeric(judge_(x,y)),
                 x = o_pick,
                 y = s_pick ) %>% t()
    return(measure = d_measures)
  }
  d_names_periods <- d_names %_% "period" %.% rep(1:number_judge_periods, each = length(d_names))
  d_periods <- lapply(1:number_judge_periods, period_compute) %>%
    do.call(rbind,.) %>%
    as_tibble(.) %>%
    magrittr::set_names("basin" %_% 1:number_judge_basins) %>%
    cbind(of = d_names_periods,.) %>%
    .[order(.$of), ]
  #
  judge_all <- rbind(d_mean,d_periods) %>%
    as_tibble(.)
  judge_all$of <- as.character(judge_all$of)
  return(judge_all)
}

```


### Bar Plots
```{r}
# ---------------------------------------------------------------------------
#' Bar plot for the Main Objective Function Values
#'
#' @rdname judge_overview
#'
#' @export
judge_barplot <- function(cosdata,
                          .ofuns = list(nse   = coscos::of_nse,
                                        kge   = coscos::of_kge,
                                        pbias = coscos::of_pbias,
                                        corr  = coscos::of_cor)
                          ) {
  # pre: ====================================================================
  cos_data <- coscos::cook_cosdata(cosdata)
  if(!(class(.ofuns) == "list")) {
    .ofuns <- list(.ofuns)
  }
  if (is.null(names(.ofuns))){
    d_names <- "d" %&% 1:length(.ofuns)
  } else {
    d_names <- names(.ofuns)
  }
  # functions: ==============================================================
  assign_ofgroups <- function(judge_melted,judge_names) {
    judge_string <- as.character(judge_melted$of)
    judge_melted$judge_group <- judge_string %>%
      replace(., startsWith(judge_string,judge_names[1]), judge_names[1]) %>%
      replace(., startsWith(judge_string,judge_names[2]), judge_names[2]) %>%
      replace(., startsWith(judge_string,judge_names[3]), judge_names[3]) %>%
      replace(., startsWith(judge_string,judge_names[4]), judge_names[4])
    return(judge_melted)
  }
  # plot-list function:
  barplot_fun <- function(judge_name,judge_melted) {
    judge_to_plot <- judge_melted %>% filter( judge_group == judge_name)
    # bad solution, but will be fine for now?
    if (judge_name == "pbias") {
      gglimits <- c(-viscos_options("of_limits")[2]*100,
                   viscos_options("of_limits")[2]*100)
    } else {
      gglimits <- viscos_options("of_limits")
    }
    plt_out <- ggplot(data = judge_to_plot) +
      geom_bar(stat = "identity",
               position = "identity",
               aes(x = of, y = value, fill = value)) +
      facet_wrap(~ variable, ncol = 1) +
      ggtitle(judge_name) +
      ylim(gglimits)
    return(plt_out)
  }
  # computations: ===========================================================
  judge_data <- judge_compute(cos_data,.ofuns)
  num_basins <- ncol(judge_data) - 1
  judge_melted <- suppressMessages( reshape2::melt(judge_data) ) %>%
    assign_ofgroups(.,d_names)
  # plotting ================================================================
  plot_list <- lapply(d_names, function(x) barplot_fun(x,judge_melted)) %>%
    magrittr::set_names(d_names)
  return(plot_list)
}

```

### Raster Plots
```{r}
#' Bar plot for the Main Objective Function Values
#'
#' @rdname judge_overview
#' @import pasta
#' @export
judge_rasterplot <- function(cosdata,
                             .ofuns = list(nse   = coscos::of_nse,
                                           kge   = coscos::of_kge,
                                           pbias = coscos::of_pbias,
                                           corr  = coscos::of_cor)
                             ) {
  # def: ====================================================================
  cos_data <- coscos::cook_cosdata(cosdata)
  if(!(class(.ofuns) == "list")) {
    .ofuns <- list(.ofuns)
  }
  if (is.null(names(.ofuns))){
    d_names <- "d" %&% 1:length(.ofuns)
  } else {
    d_names <- names(.ofuns)
  }
  # computations: ===========================================================
  regex_main_of <- d_names %.% "*"
  judge_data <- judge_compute(cos_data, .ofuns)
  #
  plot_list <- lapply(regex_main_of,function(x) plot_fun_raster(x,judge_data)) %>%
    magrittr::set_names(d_names)
  return(plot_list)
}

# plot function -------------------------------------------------------------
plot_fun_raster <- function(regex_single_of,judge_data) {
  # function definitions ====================================================
  extract_single_of <- function(judge_data){
    idx <- grep(regex_single_of,judge_data$of)
    return(judge_data[idx, ])
  }
  add_facet_info <- function(judge_data) {
    facet_column <- nrow(judge_data) %>%
      magrittr::subtract(1) %>%
      rep("period",.) %>%
      c("overall",.)
    return( cbind(judge_data,facets = facet_column) )
  }
  reverse_basin_levels <- function(prepared_data) {
    prepared_data$variable <- factor(prepared_data$variable,
                                 levels = prepared_data$variable %>%
                                   levels() %>%
                                   rev()
                                 )
    return(prepared_data)
  }
  reverse_facetting_levels <- function(prepared_data) {
    prepared_data$facets <- factor(prepared_data$facets,
                               levels = prepared_data$facets %>%
                                 levels() %>%
                                 rev()
                               )
    return(prepared_data)
  }
  bind_and_round_value <- function(of,gglimits,digits) {
    dplyr::mutate(of,
                  value = pmax(value,gglimits[1]) %>%
                    pmin(.,gglimits[2]) %>%
                    round(.,digits)
                  )
  }
  # computation =============================================================
  if (regex_single_of == "p_bias.*") {
    # pbias has different limits :(
    gglimits <- c(-viscos_options("of_limits")[2]*100,
                  viscos_options("of_limits")[2]*100)
  } else {
    gglimits <- viscos_options("of_limits")
  }
  #
  prepared_data <- judge_data %>%
    extract_single_of() %>%
    add_facet_info() %>%
    reshape2::melt(., id.vars = c("of","facets")) %>%
    reverse_basin_levels() %>%
    reverse_facetting_levels() %>%
    bind_and_round_value(.,gglimits,2)
  # ggplot ==================================================================
  plt_out <- ggplot(prepared_data,
                    aes(of,variable, fill = value),
                    environmnet = environment()) +
    geom_raster(position = "identity") +
    coord_fixed(ratio = 5)  +
    facet_grid(~ facets,  scales = "free_x", space = "free") +
    theme( legend.position = "none")  +
    geom_tile(color = "white", size = 0.25 ) +
    geom_text(aes(of,variable, label = as.character(sprintf("%0.2f",value,2))),
              color = "black")
  return(plt_out)
}

```
