# Objective Function Summaries

```{r setup6, include=FALSE, purl=FALSE}
  knitr::opts_chunk$set(eval = FALSE, tidy = FALSE)
```

This chapter presents the objective function summary functions in visCOS. They
compute basin-wise objective functions of the overall data and for the
individual basins.

As explained in [previous section](LP-distance_measures.html) objective
functions are a pivotal part of model calibration. If no specific options are
chosen visCOS uses 4 objective functions, i.e.:
$NSE$,
$KGE$,
Pearson's Correlation and
the Percentage Bias (the function definitions have been given in the chapter
about [distance metrics](LP-distance_measures.html)).

The objective functions for the overall data and the marked periods
can be computed through the function `judge_compute`. Additionally, visCOS
already provides two different options to create plots for the main objective
functions: `judge_rasterplot` and `judge_barplot`. Both functions create a list
with `ggplot` figures. Each entry in the list corresponds to one of the main
objective functions and both lists can be saved to html embedded pictures with
the [`serve` function](LP-serve.html).

The `judge` function provides a wrapper for all other `judge_` and `of_`
functions. The selection is done via the non-standardly-evaluated `type`
argument (search for lazy-evaluation if you want to know more. This implies,
among other things, that

In the following are some examples:

**Example 1.** Computing the main objective functions
```{r, message=3, purl=FALSE, eval=TRUE}
require(coscos)
require(visCOS)
require(magrittr)
#
viscos_example( ) %>%
  judge(.) %>%
  print(.)
```

**Example 2.** Plotting the results of main objective function with bar plots:
```{r, message=3, purl=FALSE, eval=TRUE}
viscos_example() %>%
  judge_barplot(.) %>%
  .[[1]] %>%
  plot(.)
```

**Example 3.** Plotting the results of main objective function with a raster:
```{r, message=3, purl=FALSE, eval=TRUE}
viscos_example() %>%
  judge_rasterplot(.) %>%
  .[[2]] %>%
  plot(.)
```

**Example 4.** Using the `judge` function as a wrapper:
```{r, message=3, purl=FALSE, eval=TRUE}
viscos_example() %>%
  judge(.,rasterplot) %>%
  .[[1]] %>%
  plot(.)
```

## Code
```{r}
# ---------------------------------------------------------------------------
# Code for the Main Objective Functions (main_of)
# authors: Daniel Klotz, Johannes Wesemann, Mathew Herrnegger
# !!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!

```

### Wrapper
```{r}
#' Wrapper for the different of functions
#'
#' This function provides a wrapper for all `judge_` and `of_` functions. The
#' input arguments `type` is lazyily-evaluated (which basically implies that
#' it does not have to be provided as a character argument).
#' which function is chosen (the "pre fix" part of
#' the selected functions, i.e. `judge_` and `of_`, do not need to be provided).
#' Th
#'
#' @import coscos
#' @import pasta
#' @import lazyeval
#' @export
judge <- function(cosdata,
                  type = "compute",
                  d_metrics = list(nse = d_nse,
                                   kge = d_kge,
                                   pbias = d_pbias,
                                   corr = d_cor),
                  opts = coscos::viscos_options(),
                  ...) {
  # lazy evaluation: ------------------------------------------------------
  le_dots <-  match.call(expand.dots = FALSE) %>%
    .["..."] %>%
    .[[1]]
  le_type <- match.call(expand.dots = FALSE) %>%
    .["type"] %>%
    .[[1]] %>% 
    as.character(.)
  if (length(le_type) == 0) {
    le_type = "compute"
  }
  if (class(le_type) != "character") {
    stop("Cannot resolve 'type' argument. Try to provide a character.")
  }
  # compute options: ------------------------------------------------------
  if (le_type == "compute") {
    judge_compute(cosdata, d_metrics)
  } else if (le_type == "judge_barplot" | le_type == "barplot") {
    judge_barplot(cosdata, d_metrics)
  } else if (le_type == "judge_rasterplot" | le_type == "rasterplot") {
    judge_rasterplot(cosdata, d_metrics)
  } else if (le_type == "of_explore" | le_type == "explore") {
    of_explore(cosdata, d_metrics)
  } else if (le_type == "of_compare" | le_type == "compare") {
    cosdata2 <- NULL
    if ("cos_data2" %in% names(le_dots)) {
      cos_data2 <- eval(le_dots$cos_data2)
    }
    of_compare(d1 = cosdata,
               d2 = cosdata2)
  } else {
    stop("There is no option (`type`) called" %&&% le_type)
  }
}
```



### Computation
The purpose of this function is to extract a the main of objective functions
from the `COSERO` data.frame. The objective functions are extracted for each
basin separately and computed for the entire length of the data, as well as
for each period separately.

The computational part of the function works as follows. In step (I) the
non-marked periods of `cos_data` (columns of
`viscos_options("name_COSperiod")` which are smaller then 0) are excluded
from further calculations. The thereby obtained data.frame is named
`evaluation_data`.
```{r}
# ---------------------------------------------------------------------------
#' Get basic objective function for cos_data
#'
#' Calculate basic objective functions(NSE, KGE, percentage BIAS, Correlation)
#' for every basin and the chosen periods.
#'
#' @param cos_data cos_data data.frame.
#' @return list of basic objective function evaluated for the different
#' hydrological years and over the whole timespan.
#'
#' @import pasta
#' @import coscos
#' @import tibble
#' @importFrom purrr map_df
#' @importFrom dplyr select filter starts_with
#' @importFrom magrittr set_names
judge_compute <- function(cosdata,
                       d_metrics = list(nse = d_nse,
                                        kge = d_kge,
                                        pbias = d_pbias,
                                        corr = d_cor),
                       opts = coscos::viscos_options()) {
  # pre: ====================================================================
  cos_data <- coscos::cook_cosdata(cosdata)
  name_period <- opts[["name_COSperiod"]]
  name_o <- opts[["name_o"]]
  name_s <- opts[["name_s"]]
  #
  if( !(class(d_metrics) == "list") ) {
    d_metrics <- list(d_metrics)
  }
  if (is.null(names(d_metrics))){
    d_names <- "d" %&% 1:length(d_metrics)
  } else {
    d_names <- names(d_metrics)
  }
  #
  evaluation_data <- cos_data[cos_data[[name_period]] > 0, ]
  number_judge_basins <- evaluation_data %>%
    names(.) %>%
    unique(.) %>%
    grepl(name_o, ., ignore.case = TRUE) %>%
    sum(.)
  data_periods <- evaluation_data %>%
    .[[name_period]] %>%
    unique(.)
  number_judge_periods <- data_periods %>% length

  # compute main-of for entire data: ========================================
  d_mean <- purrr::map_df(d_metrics, function(judge_, x, y) as.numeric(judge_(x,y)),
                 x = dplyr::select(evaluation_data,dplyr::starts_with(name_o)) %>% unname(.),
                 y = dplyr::select(evaluation_data,dplyr::starts_with(name_s)) %>% unname(.) ) %>%
    t(.) %>%
    as_tibble(.) %>%
    magrittr::set_names(., "basin" %_% 1:number_judge_basins) %>%
    cbind.data.frame(of = d_names,.)
  # compute periodwise main-of: =============================================
  period_compute <- function(k) {
    o_pick <- dplyr::filter(evaluation_data,period == data_periods[k]) %>%
      dplyr::select(.,starts_with(name_o)) %>%
      unname(.)
    s_pick <- dplyr::filter(evaluation_data,period == data_periods[k]) %>%
      dplyr::select(.,starts_with(name_s)) %>%
      unname(.)
    d_measures <- purrr::map_df(d_metrics, function(judge_,x,y) as.numeric(judge_(x,y)),
                 x = o_pick,
                 y = s_pick ) %>% t()
    return(measure = d_measures)
  }
  d_names_periods <- d_names %_% "period" %.% rep(1:number_judge_periods, each = length(d_names))
  d_periods <- lapply(1:number_judge_periods, period_compute) %>%
    do.call(rbind,.) %>%
    as_tibble(.) %>%
    magrittr::set_names("basin" %_% 1:number_judge_basins) %>%
    cbind(of = d_names_periods,.) %>%
    .[order(.$of), ]
  #
  judge_all <- rbind(d_mean,d_periods) %>%
    as_tibble(.)
  judge_all$of <- as.character(judge_all$of)
  return(judge_all)
}

```

### Plotting
```{r}
# ---------------------------------------------------------------------------
#' Plot main objective function values
#'
#' Currently two options for plotting the main objectives are provided by
#' visCOS: Plotting the different objective functions values as a set of
#' bar plots \code{barplot_of} and plotting a summary table in form of
#' a large raster of all the objective function values \code{rasterplot_of}.
#'
#' @name plot_main_of
NULL

```

#### Bar Plots
```{r}
# ---------------------------------------------------------------------------
#' Bar plot for the Main Objective Function Values
#'
#' @rdname judge_overview
#'
#' @import coscos
#'
#' @export
judge_barplot <- function(cosdata, d_metrics = list(nse = d_nse,
                                        kge = d_kge,
                                        pbias = d_pbias,
                                        corr = d_cor)) {
  # pre: ====================================================================
  cos_data <- coscos::cook_cosdata(cosdata)
  if(!(class(d_metrics) == "list")) {
    d_metrics <- list(d_metrics)
  }
  if (is.null(names(d_metrics))){
    d_names <- "d" %&% 1:length(d_metrics)
  } else {
    d_names <- names(d_metrics)
  }
  # functions: ==============================================================
  assign_ofgroups <- function(judge_melted,judge_names) {
    judge_string <- as.character(judge_melted$of)
    judge_melted$judge_group <- judge_string %>%
      replace(., startsWith(judge_string,judge_names[1]), judge_names[1]) %>%
      replace(., startsWith(judge_string,judge_names[2]), judge_names[2]) %>%
      replace(., startsWith(judge_string,judge_names[3]), judge_names[3]) %>%
      replace(., startsWith(judge_string,judge_names[4]), judge_names[4])
    return(judge_melted)
  }
  # plot-list function:
  barplot_fun <- function(judge_name,judge_melted) {
    judge_to_plot <- judge_melted %>% filter( judge_group == judge_name)
    # bad solution, but will be fine for now?
    if (judge_name == "pbias") {
      gglimits <- c(-viscos_options("of_limits")[2]*100,
                   viscos_options("of_limits")[2]*100)
    } else {
      gglimits <- viscos_options("of_limits")
    }
    plt_out <- ggplot(data = judge_to_plot) +
      geom_bar(stat = "identity",
               position = "identity",
               aes(x = of, y = value, fill = value)) +
      facet_wrap(~ variable, ncol = 1) +
      ggtitle(judge_name) +
      ylim(gglimits)
    return(plt_out)
  }
  # computations: ===========================================================
  judge_data <- judge_compute(cos_data,d_metrics)
  num_basins <- ncol(judge_data) - 1
  judge_melted <- suppressMessages( reshape2::melt(judge_data) ) %>%
    assign_ofgroups(.,d_names)
  # plotting ================================================================
  plot_list <- lapply(d_names, function(x) barplot_fun(x,judge_melted)) %>%
    magrittr::set_names(d_names)
  return(plot_list)
}

```

#### Raster Plots
```{r}
#' Bar plot for the Main Objective Function Values
#'
#' @rdname judge_overview
#' @import pasta
#' @export
judge_rasterplot <- function(cos_data, d_metrics = list(nse = d_nse,
                                        kge = d_kge,
                                        pbias = d_pbias,
                                        corr = d_cor)) {
  # def: ====================================================================
  build_tibble(cos_data)
  if(!(class(d_metrics) == "list")) {
    d_metrics <- list(d_metrics)
  }
  if (is.null(names(d_metrics))){
    d_names <- "d" %&% 1:length(d_metrics)
  } else {
    d_names <- names(d_metrics)
  }
  # computations: ===========================================================
  regex_main_of <- d_names %.% "*"
  judge_data <- judge_compute(cos_data, d_metrics)
  #
  plot_list <- lapply(regex_main_of,function(x) plot_fun_raster(x,judge_data)) %>%
    magrittr::set_names(d_names)
  return(plot_list)
}

# plot function -------------------------------------------------------------
plot_fun_raster <- function(regex_single_of,judge_data) {
  # function definitions ====================================================
  extract_single_of <- function(judge_data){
    idx <- grep(regex_single_of,judge_data$of)
    return(judge_data[idx, ])
  }
  add_facet_info <- function(judge_data) {
    facet_column <- nrow(judge_data) %>%
      magrittr::subtract(1) %>%
      rep("period",.) %>%
      c("overall",.)
    return( cbind(judge_data,facets = facet_column) )
  }
  reverse_basin_levels <- function(prepared_data) {
    prepared_data$variable <- factor(prepared_data$variable,
                                 levels = prepared_data$variable %>%
                                   levels() %>%
                                   rev()
                                 )
    return(prepared_data)
  }
  reverse_facetting_levels <- function(prepared_data) {
    prepared_data$facets <- factor(prepared_data$facets,
                               levels = prepared_data$facets %>%
                                 levels() %>%
                                 rev()
                               )
    return(prepared_data)
  }
  bind_and_round_value <- function(of,gglimits,digits) {
    dplyr::mutate(of,
                  value = pmax(value,gglimits[1]) %>%
                    pmin(.,gglimits[2]) %>%
                    round(.,digits)
                  )
  }
  # computation =============================================================
  if (regex_single_of == "p_bias.*") {
    # pbias has different limits :(
    gglimits <- c(-viscos_options("of_limits")[2]*100,
                  viscos_options("of_limits")[2]*100)
  } else {
    gglimits <- viscos_options("of_limits")
  }
  #
  prepared_data <- judge_data %>%
    extract_single_of() %>%
    add_facet_info() %>%
    reshape2::melt(., id.vars = c("of","facets")) %>%
    reverse_basin_levels() %>%
    reverse_facetting_levels() %>%
    bind_and_round_value(.,gglimits,2)
  # ggplot ==================================================================
  plt_out <- ggplot(prepared_data,
                    aes(of,variable, fill = value),
                    environmnet = environment()) +
    geom_raster(position = "identity") +
    coord_fixed(ratio = 5)  +
    facet_grid(~ facets,  scales = "free_x", space = "free") +
    theme( legend.position = "none")  +
    geom_tile(color = "white", size = 0.25 ) +
    geom_text(aes(of,variable, label = as.character(sprintf("%0.2f",value,2))),
              color = "black")
  return(plt_out)
}

```
