# Runoff Peaks

```{r setup8, include=FALSE, purl=FALSE}
  knitr::opts_chunk$set(eval = FALSE, tidy = FALSE, fig.align = 'center')
```

The function `peack` lets users explore the highest events in
among the available basins. It provides a list of [ggplot2](http://ggplot2.org/)
plots, containing an overview plot (`overview`), a scatter plot (`scatter`)
and detail plots of the individual events (`event_plot`). Instead of
explaining the properties of each plot in detail it is best to get an intuition
of the function by looking at some examples.

## Examples:
For the examples 10 events are extracted from a runoff example
```{r, eval = TRUE, purl=FALSE}
  library(coscos)
  library(visCOS)
  cosdata <- viscos_example()
  peakplots <- peack(cosdata, action = "plots", n_events = 10L)

```

The `peakplots` list does now contain plots for each basin within the
`cosdata` data.frame:
```{r, eval = TRUE, purl=FALSE}
  names(peakplots)
```

For each basin the a set of plots (`overview`,`scatter`,`event_plot`) are saved
within a list for each basin. In the following the plots for basin 1 are shown:
```{r, eval = TRUE, purl=FALSE}
  names(peakplots$basin0001)
```

The `overview` plot shows the entire time series of `data1` and `data2` of the
basin. The found events are marked with black dots. The `overview` plot for
basin 1 is:
```{r, eval = TRUE, purl=FALSE}
  peakplots$basin0001$overview
```

The `scatter` plot shows the found events within a scatter plot, where  `data1`
is the x-axis and `data2` on the y-axis. In the following an example for basin 1
is given.
```{r, eval = TRUE, purl=FALSE}
  peakplots$basin0001$scatter
```

Detail plots for each of the found events are given in form of the `event_plot`
objects. Here an example:
```{r, eval = TRUE, purl=FALSE}
  peakplots$basin0001$event_plot5
```


## Code
This part of the document defines the code of `peack`
```{r}
  # --------------------------------------------------------------------------
  #' Event-wise Exploration of the high-runoffs
  #'
  #' @export
  #'
  #' @import ggplot2
  #' @import dplyr
  #' @import magrittr
  #' @importFrom tibble tibble
  peack <- function(cosdata,
                    action = "compute",
                    n_events= 10L,
                    window_size = 24L,
                    opts = coscos::viscos_options()) {
    # pre:
    action_string <- match.call(expand.dots = FALSE) %>%
      .["action"] %>%
      .[[1]] %>%
      as.character(.)
    if (length(action_string) == 0) {
      action_string = "compute"
    }
    #
    cos_data <- coscos::cook_cosdata(cosdata)
    n_events_int <- as.integer(n_events)
    window_size_int <- as.integer(window_size)
    if (is.na(n_events_int)  |
        is.nan(n_events_int) |
        is.infinite(n_events_int) |
        !is.integer(n_events_int) ) {
      stop("n_events is ill defined")
    }
    if (is.na(window_size)  |
        is.nan(window_size) |
        is.infinite(window_size) |
        !is.integer(window_size) ) {
      stop("window_size is ill defined")
    }
    data1 <- cos_data %>%
      dplyr::select( starts_with(opts[["name_o"]]) )
    data2 <- cos_data %>%
      dplyr::select( starts_with(opts[["name_s"]]) )
    data_posix <- cos_data %>%
      dplyr::select(matches(opts[["name_COSposix"]]))
    data_numbers <- names(data1) %>%
      gsub(opts[["name_o"]], "", ., ignore.case = TRUE) %>%
      gsub("\\D","",.,ignore.case = TRUE)
    result_list <- lapply(1:ncol(data1),
                   function(x) peack_one_basin(data1[ ,x],
                                                  data2[ ,x],
                                                  data_posix,
                                                  n_events_int,
                                                  window_size_int,
                                                  action_string,
                                                  opts)) %>%
      set_names(., paste("basin", data_numbers, sep = ""))
    return(result_list)
  }
```

### Peak-Plots for a signle Basin
This is the function for generating the different plots for one basin.
At first the provided time series are grouped into a `tibble`, then the
peaks of the observations are obtained via the peak_finder function and
organized. Then `ggplot2` is used for plotting.
```{r}
  peack_one_basin <- function(qobs,
                              qsim,
                              data_posix,
                              n_events_int,
                              window_size_int,
                              action_string,
                              opts) {
    single_data <- cbind(data_posix,qobs,qsim) %>%
      magrittr::set_names(c("time","obs","sim"))
    #
    peak_idx <- find_peaks(single_data$obs,m = window_size_int)
    peak_organised <- tibble::tibble(idx = as.integer(peak_idx),
                                     time = single_data$time[peak_idx],
                                     peak_obs = single_data$obs[peak_idx],
                                     peak_sim = single_data$sim[peak_idx])
    highest_peaks_organised <- peak_organised$peak_obs %>%
      sort(decreasing = TRUE) %>%
      .[1:n_events_int] %>%
      match(.,peak_organised$peak_obs) %>%
      peak_organised[., ]
    #
    if (action_string == "compute") {
      return(highest_peaks_organised)
    } else if (action_string == "plots") {
    overview_plot <- ggplot( ) +
      geom_line(data = single_data, aes(x = time, y = sim),
                col = opts[["color_s"]]) +
      geom_line(data = single_data, aes(x = time, y = obs),
                col = opts[["color_o"]]) +
      geom_point(data = highest_peaks_organised, aes(time, peak_obs))
    overview_scatter <- ggplot() +
      geom_point(data = single_data, aes(obs,sim), color = "#DDDDDD") +
      geom_abline() +
      geom_point(data = highest_peaks_organised, aes(peak_obs,peak_sim), size = 4) +
      expand_limits(x = 0, y = 0)
    sub_plots <- lapply(
      1:nrow(highest_peaks_organised),
      function(x) sub_peakplot_fun(x,
                                   window_size_int,
                                   highest_peaks_organised[x,]$idx,
                                   single_data,
                                   opts)
      ) %>%
      set_names(.,paste("event_plot", 1:length(.), sep = ""))
    return(overview = append(list(overview = overview_plot,
                                  scatter = overview_scatter), sub_plots))
    } else {
      stop("The chosen 'action' does not exist.")
    }

  }
```

### Peak Finding Function
The function for finding the peaks was proposed and developed by the cross validated user *"stas g"*  [here](http://stats.stackexchange.com/questions/22974/how-to-find-local-peaks-valleys-in-a-series-of-data)
thread.  
This is by far not the only option/possibility to approach the peak finding task.
Other nice ideas for finding peaks can be found in [this](http://stats.stackexchange.com/questions/36309/how-do-i-find-peaks-in-a-dataset)
cross validated thread.

```{r}
  ####
  # peak finder function:
  find_peaks <- function(x, m = 3){
    shape <- diff(sign(diff(x, na.pad = FALSE)))
    pks <- sapply(which(shape < 0), FUN = function(i){
      z <- i - m + 1
      z <- ifelse(z > 0, z, 1)
      w <- i + m + 1
      w <- ifelse(w < length(x), w, length(x))
      if(all(x[c(z : i, (i + 2) : w)] <= x[i + 1])) return(i + 1) else return(numeric(0))
    })
    pks <- unlist(pks)
    pks
  }
```

### Subplot Function
This function is a wrapper around `ggplot`, which is used to generate the
individual event plots.
```{r}
  ####
  # sub plot function:
  sub_peakplot_fun <- function(x, window_size, idx, peak_data, opts) {
    point <- peak_data[idx, ]
    plot_sub <- peak_data[(idx - window_size):(idx + window_size), ] %>%
      ggplot(data = .) +
      geom_line(aes(x = time, y = sim),
                col = opts[["color_s"]]) +
      geom_line(aes(x = time, y = obs),
                col = opts[["color_o"]]) +
      geom_point(data = point, aes(time, obs))
    return(plot_sub)
  }

```

## References
- http://stats.stackexchange.com/questions/22974/how-to-find-local-peaks-valleys-in-a-series-of-data (checked 12/2016)
- http://stats.stackexchange.com/questions/36309/how-do-i-find-peaks-in-a-dataset (checked 12/2016)
