---
title: "Examples for Cooking Data"
author: "Daniel Klotz"
date: "6 Juli 2016"
output:
  html_document:
    number_sections: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Here examples for **visCOS**-functions are given that deal with the process of
transforming  *raw data* into prepared data for **visCOS**.
The data is then *cooked* and can be used for figures etc.

# Getting raw data
From the point of view of **visCOS** raw-data are time-series of observations
and model output. Usually raw data is saved in some simple file format,
e.g. *.txt* or *.csv*. `R` already includes many option to read those files
with the `read.table` functionalities.
For larger unstructured files the read options of the `data.table` package is 
recommended. In our tests it was fastest and most flexible option.
Alternatively the `readr` can be used if the data is more structured and large.

The function `get_runoff_example` can be used to get some exemplary data
from within **visCOS**:
```{r}
  library(visCOS)
  runoff_example <- get_runoff_example()
  head(runoff_example)
```
