---
title: "Aggregate and Plot"
author: 
- "Daniel Klotz, Johannes Wesemann"
date: "22 September 2016"
output:
  html_document:
    number_sections: yes
---



```{r setup, include=FALSE, purl=FALSE}
  knitr::opts_chunk$set(eval = FALSE, tidy = FALSE, fig.align = 'center')
```

# Introduction
 In hydrology it is often usefull 
to summarise the runoff with respect to a given time dimension. 
In **visCOS** this can be done by using the `aggreate_and_plot` functions. 
This is still an experimental function with a rather badly name. 
The function takes [COSERO data.frame](cook_data.html),  aggregates them 
accodring to a single or more given time resolutions and returns a `ggplot` of 
`data1` and `data2`. The different basins are faceted into different plots. 


```{r, eval = TRUE, purl=FALSE}
  runoff_data <- visCOS::get_runoff_example()
```

```{r, eval = TRUE, purl=FALSE}
  visCOS::aggregate_and_plot( runoff_data, "dd")
```

```{r, eval = TRUE, purl=FALSE}
  visCOS::aggregate_and_plot( runoff_data, "mm")
```

```{r, eval = TRUE, purl=FALSE}
  visCOS::aggregate_and_plot( runoff_data, "yyyy")
```

```{r, eval = TRUE, purl=FALSE}
  visCOS::aggregate_and_plot( runoff_data, "yyyy-mm") + 
    theme( axis.text.x = element_text(angle = 90) )
```

# Code

```{r}
  #' Aggreage and plot
  #' 
  #' Hopefully a good description will follow 
  #'
  #' @import magrittr
  #' @import ggplot2
  #' @import pasta
  
  #' @export
  aggregate_and_plot <- function( runoff_data,aggregation = "mm") {
```
The time aggregation is done by cutting the needed information out of the 
date-string. This is a bit ruff, but works nicely and seems to be more 
commonly used than you would expect. 
```{r}
    #
    cutting_bounds <- c(Inf,-Inf)
    if (grepl("dd",aggregation)) {
      cutting_bounds[1] <- min(9,cutting_bounds[1])
      cutting_bounds[2] <- max(11,cutting_bounds[2])
    }
    if (grepl("mm",aggregation)) {
      cutting_bounds[1] <- min(6,cutting_bounds[1])
      cutting_bounds[2] <- max(7,cutting_bounds[2])
    }
    if (grepl("yyyy",aggregation)) {
      cutting_bounds[1] <- min(1,cutting_bounds[1])
      cutting_bounds[2] <- max(4,cutting_bounds[2])
    }
```

```{r}
    ######  define helpers
    # Paste funcitons:
    regex_for_runoff_selection <- viscos_options("name_data1") %|%  viscos_options("name_data2")
    # aggregation function:
    aggregator_fun <- function(k,data_frame){
      the_aggregation <- aggregate(data_frame[[k]] ~ data_frame$date_selection, FUN=median)
      return(the_aggregation[ ,2])
    }
    ##### main code 
    runoff_data %>% 
      visCOS::prepare_complete_date() %>% 
      visCOS::remove_chunk() -> full_runoff_data 
    runoff_with_aggreggation <- cbind.data.frame(
      full_runoff_data ,
      date_selection = substr(full_runoff_data$posixdate,
                              cutting_bounds[1],
                              cutting_bounds[2]) %>% as.factor()
      )
    names_runoff_selection <- grep(
      regex_for_runoff_selection,
      names(runoff_with_aggreggation) %>% tolower, 
      value = TRUE
      )
    selected_runoff_rows <- grep(regex_for_runoff_selection,
                             names(runoff_with_aggreggation) %>% tolower)
    sub_selection <- sapply(selected_runoff_rows, 
                            function(x) aggregator_fun(x,runoff_with_aggreggation)) %>% 
      data.frame(.,
                 timestep = 1:nrow(.), 
                 month = unique(runoff_with_aggreggation$date_selection) ) %>% 
      set_names(.,
                c(names_runoff_selection,"timestep","month"))
    # 
    melted_sub_selection <- sub_selection %>% 
      reshape2::melt(., id.vars = c("timestep","month")) %>% 
      cbind.data.frame(., 
                       basin =  .$variable %>%
                         gsub(regex_for_runoff_selection,"",.) %>% 
                         gsub("\\D","",.) %>%
                         as.integer, 
                       obs_sim = .$variable %>% 
                         gsub(viscos_options("name_data1") %&% ".*",viscos_options("name_data1"),.) %>% 
                         gsub(viscos_options("name_data2") %&% ".*",viscos_options("name_data2"),.))
    the_plot <- ggplot(melted_sub_selection) + 
      geom_line(aes(x = timestep, y= value, col = obs_sim)) + 
      scale_colour_manual(values = c(viscos_options("color_data1"),viscos_options("color_data2"))) + 
      scale_x_discrete(limits = melted_sub_selection$month, 
                       labels = abbreviate) + 
      facet_wrap( ~ basin,
                  ncol = 2, 
                  scales = "free") +
      theme(panel.spacing = unit(1.5, "lines")) 
    return(the_plot)
    }
```



