---
title: "Code for Cooking Data"
author: 
- "Daniel Klotz, Johannes Wesemann"
date: "6 Juli 2016"
output:
  html_document:
    number_sections: yes
---

```{r setup, include=FALSE, purl=FALSE}
  knitr::opts_chunk$set(eval = FALSE, tidy = FALSE, fig.align = 'center')
```

# Introduction
This section defines code of the **visCOS** functions used for "cooking data".
Within the scope of **visCOS** "cooking data" is defined as the process of
transforming *raw data* into *cooked*, which can be used for the internal
visualisations.



# Code
The following functions are defined below. Exported functions can be seen by the
user.

function | exported
------------- | -------------
`get_runoff_example` | yes
`remove_chunk` | yes
`only_observed_basins` | no
`prepare_complete_date` | yes
`implode_cosdate` | no
`remove_leading_zeros` | no
`mark_periods` | yes
`runoff_as_xts` | no
`extract_objective_functions` | yes

## `get_runoff_example`
visCOS porvides some exemplary data. All the available functions can be tested
with it. `get_runoff_example` is a small wrapper function to get the data from
within the package via `read.csv`

**Note** at this stage, only a comparison between **numbered catchments** 
is possible for two parameters is possible. So the data must include an integer 
number after the description (like "QObs_001" and "QSim_001")
```{r, echo=FALSE}
  #' Get runoff example
  #'
  #' Get exemplary runoff data to test the different functions of visCOS
  #' @export
  
```

```{r}
  get_runoff_example <- function() {
    file_path <- system.file("extdata",
                             "runoff_example.csv",
                             package = "visCOS")
    runoff_example <- read.csv(file_path)
    return(runoff_example)
  }
```


## `remove_chunk`
This function removes all columns not specified in the `viscos_options`,
as well basins where no observations are available.
The first part is done by identifying the names of the data_frame and
searching for the idx of the names defined in the `viscos_options`.
The second part is done via the function `prepare.only_observed` (see next 
subchapter). It is actually so pretty simple, but really useful. Since the 
solution is quite specific, the function tests if the input is a data.frame and 
returns an error if not.

**Note** that the routine is **not** case sensitive. It does not distinguish
between small and capital letters!

```{r echo=FALSE}
  #' removes chunk in runoff_data
  #'
  #' Removes all columns which are not foreseen (see: viscos_options) from
  #' runoff data
  #'
  #' @param runoff_data data.frame object containing at least COSdate,
  #' and two data rows defined in viscos_options
  #' @return data.frame object without the chunk
  #' @export

```

The first part of the code loads the dependencies and makes sure that the
runoff_data variable is a data.frame (see: chapter about *defensive coding*):
```{r}
  remove_chunk <- function(runoff_data) {
    require("magrittr", quietly = TRUE)
    assert_dataframe(runoff_data)
```

The body of `remove_chunk` works as following: First, the names of the columns 
of the runoff_data are determined via `names` and all set to lower letters with
`tolower`. This is the reason why the function is not case sensitive!
Then helper function `get_regex_for_runoff_data` is called to get regular
expressions that identify the wanted columns. Later, `grep` is used to get
the index `idx` of the columns with the given names (`lowercase_names_in_data`).
Finally, `idx`is then used to select the wanted columns.

In the second part, the function `only_observed_basins` is being executed (see 
next chapter)
```{r}
  lowercase_names_in_data <- runoff_data %>% names %>% tolower
  #
  regex_columns <- get_regex_for_runoff_data() # see: helpers

  idx <- regex_columns %>%
    grep(.,lowercase_names_in_data)
    no_chunk_runoff_data <- runoff_data[ , idx]
    return( only_observed_basins(no_chunk_runoff_data) )
  }
```


### `only_observed_basins`
This function removes basins for which no observations are available. No
observation are interpreted as columns in which all the entries are either -999
or `NA's`. 
```{r, echo=FALSE}
# remove basins without observations
#
# Removes basins without observation (-999/NA values) from the provided data.frame
#
# @param runoff_data A raw runoff_data data.frame, which may contains basins
# without observations.
# \strong{Note:} It is assumed that all available basins are simulated!
# @return data.frame without the observation-free basins
# @export
```

Initiation of function: 
```{r}
only_observed_basins <- function(runoff_data) {
  require("magrittr", quietly = TRUE)
  assert_dataframe(runoff_data)
```

In order to handle the `NA's` and -999 in the same way all `NA's` are replaced
with -999. Then the `lapply` function is used to find the max of each column.
```{r}
  runoff_data[is.na(runoff_data)] <- -999
  colmax <- lapply(X = runoff_data, FUN = max) # get max of any column
```

If any column has a max of -999 it is worth further investigation.
Unfortunately the -999 columns could also mean, that something went wrong
with the simulation, thus one has to make sure that only observation columns
are selected. To do so we first identify the idx of the max == -999 cols and
build a regular expression with the help of the `viscos_options`.
We combine both information via `grepl` to check if only observed columns have
been selected. If so everything is fine, but if `any` of the selected cols came
from an simulation col the function stops the execution and gives a warning
message.
```{r}
  if ( any(colmax < 0.0) ){
    idx_temp <- which(colmax < 0.0)
    obs_regex <- paste(viscos_options("name_data1"),".*", sep ="")
    OnlyQobsSelected <- idx_temp %>%
      names %>%
      tolower %>%
      grepl(obs_regex,.) %>%
      any
    if (!OnlyQobsSelected){
      stop("There are Qsim withouth simulation (i.e. only values smaller 0). Pls remove them first")
    }
```
The next step involves a little trick in which +1 is added to the `idx` to
obtain the indexes of the corresponding simulation. Here we assume that for each
observation there is a corresponding simulation **in next the next colum**!
This allows us then to select only the observed basins. But now all NA values
are actually -999! Thus, we replace them again back to NA, via `lapply(...,min)`
assuming that all negative values are NAs. This is not a strong assumption,
because there should never be negative flows! In the end the resulting
data.frame is returned.

```{r}
    idx_slct <- c(idx_temp,idx_temp + 1) %>% sort()
    d_onlyObserved <- runoff_data[-idx_slct]
    # set remaining negative Qobs to NA, so that HydroGOF can be used correctly, also ignoring NAs
    colmin <- lapply(X = d_onlyObserved, FUN = min)
    idx_temp <- which(colmin < 0.0)
    d_onlyObserved[d_onlyObserved[idx_temp]<0, idx_temp] <- NA
    return(d_onlyObserved)
  } else {
    return(runoff_data)
  }
}


```


## `prepare_complete_date`
This function is a wrapper. It is however not finished yet!
The idea is the following. The idea is to provide a handy function to complete
the date-columns. Within the data.frame the user has to provide either 5 columns
with the *COSdate* format: `yyyy-mm-dd-hh-min` i.e.: year-month-day-hour-minute or
an equivalent column in `POSIXct` format
(see: [link](https://stat.ethz.ch/R-manual/R-devel/library/base/html/as.POSIXlt.html)).

*However*, it is currently only possible to convert *COSdate* into `POSIXct` dates
via the `implode_cosdate` function.

The function itself is basically a switch which calls different functions
if `name_cosyear` and/or `name_posix` is found and returns an error if none of
the two are found. If the user has not named his columns according to COSERO
convention he can provide the names of the *COSdate* year via `name_cosyear` and
the names of the POIXct column via `name_posix` within the `viscos_options`. The function returns an error
if `runoff_data` is not a data frame and if `OK_COSdate` or `OK_POSIXdates` are
not logical due to some problems with the dates/time-formats.

A *caveat* with this method is that only the `name_cosyear` is checked to infer
if the 5 *COSdate* columns are available. This can cause errors later if some
other columns are missing. Sadly it is not apparent how to avoid that problem,
without defining the names of all columns.

Furthermore, we need to be aware that with *UTC* a fixed time-zone is
assumed to avoid problems with leaps in time (summer/winter time).

The function `implode_cosdate` is used to obtain the POSIXct date from the
*COSdate* columns. The following chapter will show how this is done.
```{r, echo=FALSE}
#' Complete the date-formats with POSIXct or COSdate
#'
#' Complete the data-formats of your data.frame `POSIXct` and/or `COSdate`
#'
#' @param runoff_data The data.frame, which contains the runoff information
#' @param name_cosyear string with the name of the `COSdate` year column
#' @param name_posix string with the name of the POSIXct column
#' @return The new runoff data.frame with the added data-format.
#' @export
```

```{r}
prepare_complete_date <- function(runoff_data = NULL,
                                  name_cosyear = "yyyy",
                                  name_posix = "POSIXdate") {
  # make sure that magrittr is loaded:
  require("magrittr", quietly = TRUE)
  assert_dataframe(runoff_data)
  # check for COSdates and stop if non-logical expression are obtained
  OK_COSdate <- any(names(runoff_data)== viscos_options("name_COSyear"))
  OK_POSIXdates <- any(names(runoff_data)== viscos_options("name_COSposix"))
  if ( !is.logical(OK_COSdate) | !is.logical(OK_POSIXdates) ) {
    stop("Something seems to be wrong with the date / time formats :(")
  }
  # choose function depending on which formats are available!
  if (!OK_COSdate & !OK_POSIXdates) {
    stop("No COSdates and no POSIXct-dates in the data!")
  } else if (OK_COSdate & !OK_POSIXdates) {
    runoff_data <- implode_cosdate(runoff_data) # see following chapter
  } else if (!OK_COSdate & OK_POSIXdates) {
    stop("POSIXct to COSdates not yet supported :(")
  }
  return(runoff_data)
}
```

### `implode_cosdate`
This function is used to transform the old-school *COSdate* format into the
widely spread POSIXct format, used within R and in many packages. The function
is called by the function above and not provided to the user! It takes a
data frame (`runoff_data`) and uses `viscos_options` to transform the different
*COSdate* columns into POSIXct dates via `paste`.
```{r, echo=FALSE}
# transform COSdate into the nicer POSIXct-date format
#
# Takes a data.frame, which contains the COSdate format (see: xxx) and
# transforms it into a POSIXct series. Note that time is assumed to be in UTC
```
The paste function is used to concatenate the different *COSdate*
columns (i.e. year, month, day, hour and minute) to a POSIXct date format.
The obtained series of strings is saved into the variable `POSIXdate` and
added as the column `viscos_options("name_COSposix")` (see section about
`viscos_options`) to the `runoff_data.
```{r}
implode_cosdate <- function(runoff_data) {
  require("magrittr", quietly = TRUE)
  assert_dataframe(runoff_data)
  name_string <-  runoff_data %>% names %>% tolower
  #
  POSIXdate <- paste(runoff_data[[viscos_options("name_COSyear")]],
                     sprintf("%02d",runoff_data[[viscos_options("name_COSmonth")]]),
                     sprintf("%02d",runoff_data[[viscos_options("name_COSday")]]),
                     sprintf("%02d",runoff_data[[viscos_options("name_COShour")]]),
                     sprintf("%02d",runoff_data[[viscos_options("name_COSmin")]]),
                     sep= "" ) %>%
    as.POSIXct(format = "%Y%m%d%H%M",tz = "UTC")
  runoff_data[[viscos_options("name_COSposix")]] <- POSIXdate
  return(runoff_data)
}
```


## `remove_leading_zeros`
This function removes leading zeros from column names of the data.frame
`runoff_data`.The function has no defensive coding, apart from the execution
of `remove_chunk` (see corresponding chapter) to be sure that no non-needed
columns are within the data.frame.

```{r}
remove_leading_zeros <- function(runoff_data) {
  require("magrittr", quietly = TRUE)
  runoff_data %<>% remove_chunk
```

The body of the function works in the following way:
The variables `runoff_names` is computed by obtaining the names of the columns
of `runoff_data` with `names` and removing all the digits with
`gsub("\\d",...)`. The basin digits, `runoff_nums` are computed by doing the
oposite, i.e.: removing all the non digits via `gsub("\\D",...)`. Afterwards,
the empty strings ("") are replaced with a small trick: The characters are
transformed into a numerics (induces `NA's`) and back again.
In the last step the variables `runoff_names` and `runoff_nums` are simply put
togehter again via the paste command. The result has no leading zeros as in the
variable numeration.
```{r}
  runoff_names <- runoff_data %>% names %>% gsub("\\d","",.)
  # get numbers and remove leading zeros
  runoff_nums <- runoff_data %>%
    names %>%
    gsub("\\D","",.) %>%
    as.numeric %>%  
    as.character
  runoff_nums[is.na(runoff_nums)] = ""
  # paste new nums as new data_names
  names(runoff_data) <- paste(runoff_names, runoff_nums, sep = "")
  return(runoff_data)
}
```


## `mark_periods`
This function can be used to create a periods column in your data.frame. The
period column is an integer vector, which is a multiple of 1 if the row is 
within a certain period and 0 if it is out of the period.

Building it was actually quite tricky. The author himself (Klotz) is not
sure if it is really a nice solution. But it is the best solution, in terms of
speed and clearness, that we could come up so far. So take it as a challenge
to build a better solution!

As shown below the function takes the `runoff_data` data.frame and the two
integers `start_month`and `end_month`. They allow the user to define the periods
within the `runoff_data` on a monthly resolution. As the name suggests
`start_month` defines the beginning of the period and `end_month` the end of
the period. The function requires `dplyr` and `magrittr` and tests if
`runoff_data` is a data.frame. An error is returned if it is not.
Furthermore, if there is *chunk* in the data.frame it is removed by
`remove_chunk` and `prepare_complete_date` (see above)is used to ensure that
the full time-information, i.e. both date formats, is available.
```{r, echo=FALSE}
#' calculate periods
#'
#' Mark the periods within runoff_data.
# The marking uses a monthly resolution, which are defined by the integers
#' `start_month` and `end_month`.  
#'
#' @param runoff_data The data.frame, which contains the runoff information
#' @return The runoff data.frame reduced and ordered according to the
#' hydrological years within the data.
#' \strong{Note:} The periods columns are formatted as characters!
#' @export
```

```{r}
mark_periods <- function(runoff_data, start_month = 10, end_month = 9) {
  require("dplyr", quietly = TRUE, warn.conflicts = FALSE)
  require("magrittr", quietly = TRUE)
  assert_dataframe(runoff_data)
  runoff_data %<>% remove_chunk %>% prepare_complete_date()

```


The body of the function is split in two parts.
In the *first part* the variable `period_range` is defined as a integer sequence
from the `start_month` to the `end_month`.
The solution distinguishes between two different cases:
`start_month` <= `end_month` and `start_month` > `end_month`. The solution
for the former is trivial, as `seq` can be used. In principle the solution for 
the latter is also simple as two ranges (`range_1` and `range_2`) can be
constructed and concatenated. Note that both solutions additionally construct 
the integer-vector `out_of_period`. The vector contains information about the 
months which are not within the period_range. It would not be necessary for this 
case. But, since the current solution still seems to be suboptimal, it is 
needed for the second case, in order to mark the months which are out of the 
period later.
```{r}
  # (I) get labels for the months
  if (start_month <= end_month ) {
    period_range <- seq(start_month,end_month)
    out_of_period <- seq(1,12) %>% extract( !(seq(1,12) %in% period_range) )
  } else if (start_month > end_month) {
    range_1 <- seq(start_month,12)
    range_2 <- seq(1,end_month)
    period_range <- c(range_1,range_2)
    out_of_period <- seq(1,12) %>% extract( !(seq(1,12) %in% period_range) )
  }
```

The *second part* is a bit trickier.

First, the function `eval_diff` is defined. It takes a vector `a` and 
calculates the differences for it, where the first entry of `a` is kept,
e.g. `eval_diff( c(3,2,5,7,8) )` is `3 -1  3  2  1`.
Then, `eval_diff` is used to to mark all the starting months of the runoff data 
and saves them into a `period` column. The `pmax(.,0)` operation guarantees that
negative differencing (as above in 3 -> 2) is not included and the cumulative
sum is used to count the periods are within the data.frame. Therefore, at the 
beginning of the first period, the counter jumps to "1" and becomes "2" with the 
beginning of the second period and so on. 

Secondly, the `out_of_period` of all years is set back to zero again by simply
checking which months of our data.frame are equal to the `out_of_period` entries.
There are two problems with that solution. One is that the last year is not
extracted properly if `start_month` > `end_month`. For this case `dplyr::mutate`
is used to set the out-of-period-dates from the last year to 0, if they are 
bigger then `end_month`.
The second problem is that the first and last period are also included in the
solution even if they are **not** complete! *Thus, if a user does not want 
these solutions to be "marked" he has to remove it himself!*
```{r}
  # (II) mark periods:
    eval_diff <- function(a) {c(a[1],diff(a))}
    runoff_data[[viscos_options("name_COSperiod")]] <- 
      runoff_data[[viscos_options("name_COSmonth")]] %in% c(start_month) %>%
      eval_diff %>%
      pmax(.,0) %>%
      cumsum
    runoff_data$period[runoff_data[[viscos_options("name_COSmonth")]] %in% out_of_period] <- 0
    # corrections for last year
    max_year <- max(runoff_data[[viscos_options("name_COSyear")]])
    runoff_data %<>% dplyr::mutate(
      period = ifelse(
        (  (.[[viscos_options("name_COSyear")]] == max_year) &
           (.[[viscos_options("name_COSmonth")]] > end_month)  ),
                      0,
                      period
        )
      )
    return(runoff_data)
}
```
Here are two example plots to display the periods numbering:
```{r, eval=TRUE, purl=FALSE}
require(magrittr)
require(visCOS)
# mark_periods example 1: Hydrological years (September till August)
# note that the last year is not complete, so the counter jumps back to 0
ex1 <- get_runoff_example() %>% mark_periods(.,start_month = 9, end_month = 8)
plot(ex1$period, xlab="Timestep", ylab="# of hydrological year")

# mark_periods example 2: Summer Months (June till August)
ex2 <- get_runoff_example() %>% mark_periods(.,start_month = 6, end_month = 8)
plot(ex2$period, xlab="Timestep", ylab="# year of summer months")

```



## `runoff_as_xts`
This function is actually just a small wrapper around the `xts::xts` function 
for the purposes of **visCOS**. `runoff_as_xts` depends on `magrittr`, `xts` and
`zoo` (where the last one is acutally a dependency of `xts`).
The function takes the runoff_data as input and makes sure that the data is a
data.frame (check: `assert_dataframe` in the **defensive code** chapter),
has no "junk" inside (`assert_chunk`, see also: `remove_chunk`) and
has all the needed date/time columns (`assert_complete_date`, see also:
`prepare_complete_date`).
```{r, echo=FALSE}
# Convert runoff_data to xts-format
#
# Converts the runoff_data (class: data_frame) into an xts object
#
# @param runoff_data data_frame of the runoff_data (see: xxx)
# @return xts object of the runoff_data data.frame
# @export
```
```{r}
runoff_as_xts <- function(runoff_data) {
  # pre
  require("zoo", quietly = TRUE, warn.conflicts = FALSE)
  require("xts", quietly = TRUE, warn.conflicts = FALSE)
  require("magrittr", quietly = TRUE)
  assert_dataframe(runoff_data)
  assert_chunk(runoff_data)
  assert_complete_date(runoff_data)
```

A notable quirk of the function is that the names of the header are put to lower
cases via the `tolower` function and possible leading zeros in the enumeration
of the basins are removed.

Currently it is not exported as users can use `xts` themselves perfectly well,
and it is felt that the function does not provide enough added value for the
user.

```{r}
  # everything is set tolower because we try to keep visCOS case insensitive
  runoff_data <- remove_leading_zeros(runoff_data)
  names(runoff_data) <- names(runoff_data) %>% tolower
  name_posix <- viscos_options("name_COSposix") %>% tolower
  runoff_data_as_xts <- xts::xts(x = runoff_data,
                                 order.by = runoff_data[[name_posix]])
  #
  return(runoff_data_as_xts)
}
```
