# Flow Duration Curves

```{r setup4, include=FALSE, purl=FALSE}
  knitr::opts_chunk$set(eval = FALSE, tidy = FALSE)
```

Flow-duration curves represent the relationship between the magnitue and the
frequency of a streamflow. They provide an estimate of the percentage of time
a given streamflow was exceeded within the evaluated time frame. Foster [1934]
attributes the first use of flow duration curves to Clemens Herschel around 1880.
Since they have been used for a wide array of applications.
**visCOS** provides a function to compute the data for flow-duration curves and
a function to plot them directly.
The former function is called `fdc_compute`. It computes the flow exceedance
properties and returns a data.frame. The calculations are adapted from the
method used within the [hydroTSM](https://CRAN.R-project.org/package=hydroTSM)
package. It is currently rather slow.
The plot-function is called `fdc_plot`. Internally it uses `fdc_compute` for the
data preparation and generates a facetted `ggplot` object from it. In the plot
each basin is a facet and each sub-plot shows the $o$-data and $s$-data (see:
[Introduction](Introduction.html)).

## Example
Flow duration curves can be plotted in **visCOS** in the following way:
```{r example, purl=FALSE, eval=TRUE}
  library(visCOS)
  cos_data <- get_cos_data_example()
  fdc_plot(cos_data)
```

## Code
### Computing the data for the flow duration curves
```{r}
  #' Compute Flow Duration Curves
  #'
  #' Computes the flow duration curves (fdc) for the `cos_data` data.frame.
  #' The calculations are adapted from the method used within the hydroTSM package.
  #' @param cos_data A data.frame with columns as used throughout visCOS
  #' @import magrittr
  #' @import dplyr
  #' @importFrom purrr map_df
  #' @import pasta
  #' @export
  fdc_compute <- function(cos_data) {
    # defensive code:
    assert_dataframe(cos_data)
    # calculations:
    cos_data_only <- cos_data %>%
      select(starts_with(viscos_options("name_o")), starts_with(viscos_options("name_s")))
    cos_exceedances <- map_df(cos_data_only,calc_percent_exceedance)
    fdc_data <- cos_data_only %>% tidyr::gather() %>%
      cbind.data.frame(exceedance = cos_exceedances %>%
                         tidyr::gather() %>%
                         magrittr::extract("value")
                       ) %>%
      magrittr::set_names(c("key","value","exceedance")) %>%
      mutate(obs_sim = key %>%
               gsub( viscos_options("name_o") %&% ".*",
                     viscos_options("name_o"),
                     .,
                     ignore.case = TRUE ) %>%
               gsub( viscos_options("name_s") %&% ".*",
                     viscos_options("name_s"),
                     .,
                     ignore.case = TRUE ),
             basin_idx = key %>%
               gsub(viscos_options("name_o"),"",.,ignore.case = TRUE) %>%
               gsub(viscos_options("name_s"),"",.,ignore.case = TRUE) %>%
               gsub("\\D","",.) %>% as.numeric)
    return(fdc_data)
  }
```

```{r}
# function to calculated the percent exceedance (x-axis) for the fdc
  calc_percent_exceedance <- function(q) {
    q_sorted <- sort(q)
    q_zero_index <- which(q_sorted == 0)
    nzeros <- length(q_zero_index)
    ind <- match(q, q_sorted)
    n <- length(q)
    percent_exeedence <- rep(NA, n)
    percent_exeedence[1:n] <- sapply(1:n, function(j, y) {percent_exeedence[j] <- length(which(y >= y[j]))},
                      y = q)
    percent_exeedence <- percent_exeedence/n
    return(percent_exeedence)
  }
```

### Plotting the Flow duration curves
```{r}
  #' Plot Flow Duration Curves
  #'
  #' Plots the flow duration curves (fdc) for `cos_data`.
  #' The function uses `ggplot` to so and facets the different basins into
  #' separate subplots. Each subplot shows the fdc of the \eqn{o}-data and
  #' the \eqn{s}-data.
  #' @export
  #' @import ggplot2
  fdc_plot <- function(cos_data,
                       log_y = TRUE,
                       log_x = FALSE,
                       ...) {
  # maybe we have to account certain limits for the logs
  # if (log_y | log_x & min(ylim) == 0) {
  #   ylim <- range(q, na.rm = TRUE)
  #   tmp <- unlist(q)
  #   tmp[which(tmp == 0)] <- NA
  #   ylim[1] <- min(tmp, na.rm = TRUE)
  # }
  fdc_data <- fdc_compute(cos_data)
  gplot <- ggplot(fdc_data)
  if (log_x & log_y) {
    gplot <- gplot + geom_line(aes(x = log(exceedance), y = log(value), color = obs_sim))
  } else if (log_y) {
    gplot <- gplot + geom_line(aes(x = exceedance, y = log(value), color = obs_sim))
  } else if (log_x) {
    gplot <- gplot + geom_line(aes(x = log(exceedance), y = value, color = obs_sim))
  } else (
    gplot <- gplot + geom_line(aes(x = exceedance, y = value, color = obs_sim))
  )
  gplot <- gplot + facet_wrap(~ basin_idx)
  return(gplot)
  }
```

## References
- Foster, H.A. (1934): Duration curves. ASCE Trans., 99, 1213-1267
- Vogel, R. M.; Fennessy, N. M. (1994): Flow-Duration Curves. I: New Interpretation
and Confidence Intervals. JWRMD 120, No. 4
