---
title: "helpers"
author: "Daniel Klotz"
date: "8 Juli 2016"
output: 
  html_document:
    number_sections: yes
---

```{r setup, include=FALSE, purl=FALSE}
knitr::opts_chunk$set(eval = FALSE, tidy = FALSE)
```

# Introduction 
This section collects small and helpfull functions/scripts that are used 
throughout **visCOS**.

None of the functions is exported!

# Code 
The subsequent functions are defined below 

function | exported
------------- | -------------
`get_regex_for_runoff_data` | yes

## `get_regex_for_runoff_data`
This function can be called to get the names of the 8 allowed column-names 
within **visCOS**. `get_regex_for_runoff_data` takes no input, as it gets the 
information directly from the global options (see: *viscos_options section*). 
```{r}
get_regex_for_runoff_data <- function() {
  regex_pattern <- paste("^",viscos_options()$name_COSyear,"$|", 
                         "^",viscos_options()$name_COSmonth,"$|",
                         "^",viscos_options()$name_COSday,"$|",
                         "^",viscos_options()$name_COShour,"$|",
                         "^",viscos_options()$name_COSmin,"$|",
                         viscos_options()$name_COSobs,".*|",
                         viscos_options()$name_COSsim,".*|",
                         viscos_options()$name_COSposix,"|",
                         viscos_options()$name_COSperiod,
                         sep = "")
  return(regex_pattern)
}
```