# Summaries of important objective functions

```{r setup6, include=FALSE, purl=FALSE}
  knitr::opts_chunk$set(eval = FALSE, tidy = FALSE)
```

This chapter explains the code to calculate the main objective functions $of$
used in visCOS. As explained in [respective section](LP-of.html) objective
functions are a pivotal part of model calibration. As of now, **visCOS** focuses
on 4 main objective function: NSE, KGE, Pearson's Correlation and the Percentage
bias (the respective definitions are given [here](LP-of.html)).

The main objective functions for the overall data and the marked periods
can be computed through the function `main_of_compute`. In order to run the
function the period have to be marked first, e.g. through the `mark_periods`
function. Additionally, **visCOS** already provides two different options
to create plots for the main objective functions: `main_of_rasterplot`
and `main_of_barplot`. Both functions create a list with 4 `ggplot` figures. Each entry
in the list corresponds to one of the main objective functions and both lists
can be saved to html embedded .jpgs with the [`serve` function](LP-serve.html).

**Examples:**

Computing the main objective functions
```{r, message=3, purl=FALSE, eval=TRUE}
require(visCOS)
require(magrittr)
of_values <- get_cos_data_example() %>%
  mark_periods() %>%
  main_of_compute()
of_values
```

Plotting the results of main objective function with bar plots:
```{r, message=3, purl=FALSE, eval=TRUE}
of_values <- get_cos_data_example() %>%
  mark_periods() %>%
  main_of_barplot() %>%
  extract2(1) %>%
  plot()
```

Plotting the results of main objective function with a raster:
```{r, message=3, purl=FALSE, eval=TRUE}
of_values <- get_cos_data_example() %>%
  mark_periods() %>%
  main_of_rasterplot() %>%
  extract2(2) %>%
  plot()
```

## Code
In this section the code for the `main_of` function family is defined

### Compute the Main Objective Functions
The purpose of this function is to extract a the main of objective functions
from the `COSERO` data.frame. The objective functions are extracted for each
basin separately and computed for the entire length of the data, as well as
for each period separately.

The computational part of the function works as follows. In step (I) the
non-marked periods of `cos_data` (columns of
`viscos_options("name_COSperiod")` which are smaller then 0) are excluded
from further calculations. The thereby obtained data.frame is named
`evaluation_data`.
```{r}
#' Get basic objective function for cos_data
#'
#' Calculate basic objective functions(NSE, KGE, percentage BIAS, Correlation)
#' for every basin and the chosen periods.
#'
#' @param cos_data cos_data data.frame.
#' @return list of basic objective function evaluated for the different
#' hydrological years and over the whole timespan.
#'
#' @import hydroGOF
#' @import dplyr
#'
#' @export
main_of_compute <- function(cos_data) {
  assert_dataframe(cos_data)
  if( !exists(viscos_options("name_COSperiod"), where = cos_data) ) {
    stop("Error! Period-Column missing in cos_data, use `mark_periods`")
  }
  evaluation_data <- 
    cos_data[cos_data[[viscos_options("name_COSperiod")]] > 0, ]
  number_of_basins <- evaluation_data %>%
    names %>%
    unique %>%
    grepl(viscos_options("name_o"), ., ignore.case = TRUE) %>%
    sum
  periods_in_data <- evaluation_data %>%
    magrittr::extract2(viscos_options("name_COSperiod")) %>%
    unique
  number_of_periods <- periods_in_data %>% length
  # compute main-of for entire data: ----------------------------------------
  selected_o <- dplyr::select(evaluation_data,
                              starts_with(viscos_options("name_o"))
                              ) %>% 
    unname
  selected_s <- dplyr::select(evaluation_data,
                              starts_with(viscos_options("name_s"))
                              ) %>%
    unname
  nse_ <- hydroGOF::NSE(selected_s,selected_o)
  kge_ <- hydroGOF::KGE(selected_s,selected_o)
  p_bias_ <- hydroGOF::pbias(selected_s,selected_o)
  corr_ <- cor(selected_s,selected_o) %>% diag(.)
  # compute periodwise main-of: ---------------------------------------------
  # pre allocations
  NSE_period <- matrix(nrow = number_of_periods,
                       ncol = as.integer(number_of_basins),
                       data = NA)
  KGE_period <- NSE_period
  p_bias_period <- NSE_period
  CORR_period <- NSE_period
  # calculation loop, proabbly slow :( 
  for (k in 1:number_of_periods) {
    selected_o <- dplyr::filter(evaluation_data,period == periods_in_data[k]) %>%
      dplyr::select(.,starts_with(viscos_options("name_o"))) %>%
      unname
    selected_s <- dplyr::filter(evaluation_data,period == periods_in_data[k]) %>%
      dplyr::select(.,starts_with(viscos_options("name_s"))) %>%
      unname
    NSE_period[k,1:number_of_basins] <- hydroGOF::NSE(selected_s,selected_o)
    KGE_period[k,1:number_of_basins] <- hydroGOF::KGE(selected_s,selected_o)
    p_bias_period[k,1:number_of_basins] <- hydroGOF::pbias(selected_s,selected_o)
    CORR_period[k,1:number_of_basins] <- cor(selected_s,selected_o) %>% diag(.)
  }
  # clean up: ---------------------------------------------------------------
  obj_names <- c("NSE","KGE","p_bias","CORR",
                    paste("NSE_period",1:number_of_periods,sep="."),
                    paste("KGE_period",1:number_of_periods,sep="."),
                    paste("p_bias_period",1:number_of_periods,sep="."),
                    paste("CORR_period",1:number_of_periods,sep=".")
  )
  obj_fun <- data.frame(of = obj_names,
                        basin = rbind(nse_,
                                      kge_,
                                      p_bias_,
                                      corr_,
                                      NSE_period,
                                      KGE_period,
                                      p_bias_period,
                                      CORR_period),
                        row.names = NULL)
  return(obj_fun)
}
```

### Plotting
```{r}
#' Plot main objective function values
#'
#' Currently two options for plotting the main objectives are provided by
#' visCOS: Plotting the different objective functions values as a set of
#' bar plots \code{barplot_of} and plotting a summary table in form of
#' a large raster of all the objective function values \code{rasterplot_of}.
#'
#' @name plot_main_of
NULL
```

#### Bar Plot
```{r}
#' Bar plot for the Main Objective Function Values
#'
#' @rdname of_overview
#' @export
main_of_barplot <- function(cos_data) {
  main_of_names <- c("NSE","KGE","CORR","p_bias")
  assert_dataframe(cos_data)
  of <- main_of_compute(cos_data)
  # calculate objective functions
  num_basins <- ncol(of) - 1
  # melt the of:
  # a bit of a mess, and there is probably a better solution...
  melted_of <- suppressMessages( reshape2::melt(of) ) %>%
    mutate(of_group = ifelse(of %>% as.character %>% startsWith(.,main_of_names[1]),main_of_names[1],
                             ifelse(of %>% as.character %>% startsWith(.,main_of_names[2]),main_of_names[2],
                                    ifelse(of %>% as.character %>% startsWith(.,main_of_names[3]),main_of_names[3],
                                           ifelse(of %>% as.character %>% startsWith(.,main_of_names[4]),main_of_names[4],NA)))))
  # define plot-list function
  plotlist_fun_barplot <- function(of_name) {
    of_to_plot <- melted_of %>% filter( of_group == of_name)
    if (of_name == "p_bias") {
      gglimits <- c(-viscos_options("of_limits")[2]*100,
                   viscos_options("of_limits")[2]*100)

    } else {
      gglimits <- viscos_options("of_limits")
    }
    plt_out <- ggplot(data = of_to_plot) +
      geom_bar(stat = "identity",
               position = "identity",
               aes(x = of,
                   y = value,
                   fill = value)) +
      facet_wrap(~ variable, ncol = 1) +
      ggtitle(of_name) +
      ylim(gglimits)
    return(plt_out)
  }
  # apply plot.list function to the different groupings
  plot_list <- lapply(main_of_names, plotlist_fun_barplot)
  names(plot_list) <- main_of_names
  return(plot_list)
}
```

#### Raster-Plot
```{r}
#' Bar plot for the Main Objective Function Values
#'
#' @rdname of_overview
#' @import pasta
#' @export
main_of_rasterplot <- function(cos_data) {
  main_of_names <- c("NSE","KGE","CORR","p_bias")
  regex_main_of <- main_of_names %&% ".*"
  assert_dataframe(cos_data)
  of <- main_of_compute(cos_data)
  #
  plot_list <- lapply(regex_main_of,function(x) plot_fun_raster(x,of)) %>%
    set_names(main_of_names)
  return(plot_list)
}
```

Function to plot for one basins:
```{r}
plot_fun_raster <- function(regex_single_of,of) {
  # function definitions ----------------------------------------------------
  extract_single_of <- function(of){
    idx <- grep(regex_single_of,of$of)
    return(of[idx, ])
  }
  add_facet_info <- function(of) {
    facet_column <- nrow(of) %>% 
      magrittr::subtract(1) %>% 
      rep("period",.) %>% 
      c("overall",.)
    return( cbind(of,facets = facet_column) )
  }
  reverse_basin_levels <- function(plot_data) {
    plot_data$variable <- factor(plot_data$variable,
                                 levels = plot_data$variable %>%
                                   levels() %>%
                                   rev()
                                 )
    return(plot_data)
  }
  reverse_facetting_levels <- function(plot_data) {
    plot_data$facets <- factor(plot_data$facets,
                               levels = plot_data$facets %>%
                                 levels() %>%
                                 rev()
                               )
    return(plot_data)
  }
  bind_and_round_value <- function(of,gglimits,digits) {
    dplyr::mutate(of,
                  value = pmax(value,gglimits[1]) %>% 
                    pmin(.,gglimits[2]) %>% 
                    round(.,digits)
                  )
  }
  # computation -------------------------------------------------------------
  if (regex_single_of == "p_bias.*") {
    # pbias has different limits :(
    gglimits <- c(-viscos_options("of_limits")[2]*100,
                  viscos_options("of_limits")[2]*100)
  } else {
    gglimits <- viscos_options("of_limits")
  }
  #
  plot_data <- of %>%
    extract_single_of() %>%
    add_facet_info() %>%
    reshape2::melt(., id.vars = c("of","facets")) %>%
    reverse_basin_levels() %>%
    reverse_facetting_levels() %>%
    bind_and_round_value(.,gglimits,2)
  # ggplot ------------------------------------------------------------------
  plt_out <- ggplot(plot_data,
                    aes(of,variable, fill = value),
                    environmnet = environment()) +
    geom_raster(position = "identity") +
    coord_fixed(ratio = 5)  +
    facet_grid(~ facets,  scales = "free_x", space = "free") +
    theme( legend.position = "none")  +
    geom_tile(color = "white", size = 0.25 ) +
    geom_text(aes(of,variable, label = as.character(value,2)),
              color = "black")
  return(plt_out)
}
```
