---
title: "Code for plotlist of"
author: 
- "Daniel Klotz, Johannes Wesemann, Mathew Herrnegger"
date: "14 Juli 2016"
output: html_document
---

```{r setup, include=FALSE, purl=FALSE}
  knitr::opts_chunk$set(eval = FALSE, tidy = FALSE)
```

# Code 
In this section the code of the plotlist functions are defined

## `make_of_plotlist`
The wrapper function for the different plotlists of the objective functions

```{r, echo=FALSE}
#' plotlists for the objective functions 
#'
#' To be written soon
#' @export
```

```{r}
make_plotlist <- function(runoff_data, 
                          kind_of_plot = "barplot",
                          of_name = "NSE_period") {
  if (kind_of_plot == "barplot") {
    plot_list <- serve.plotlist_periodOF(runoff_data,of_name)
  } else {
    stop("no such option for the kind of plot")
  }
  return(plot_list)
}
  

```


Users can define the to-be plotted objective function with the character 
`of_name`. 

The description for the documentation is: 


The function header is: 
```{r}
serve.plotlist_periodOF <- function(runoff_data, of_name = "NSE_period") {
  require("ggplot2", quietly = TRUE)
  require("magrittr", quietly = TRUE)
  #################
```

The first step is the calculation of the objective functions (`of`) via 
[`extract_objective_functions`](LP-computing_of.html).
Then some housekeeping needs to be undertaken. For the computations and 
selections all names are set to lower case (`tolower`), because **visCOS** works
case-insensitive. The basin enumeration `num_basins` is obtained with
the function [`get_basin_numbers`](LP-helpers.html). The plot-titles 
of the individual barplots are build by gathering the main title from the 
[global option](LP-viscos_options.html) *"plot_title"* and concatenating it with 
`num_basins`. The user defined objective function (`user_of`) is extracted from 
`of` by using the user defined selection `of_name`. Again, in order to avoid 
lower-upper case distinctions the `of_name` string is allways `set tolower`. 

```{r}
  # calculate objective functions
  of <- extract_objective_functions(runoff_data)
  # 
  names(of) <- tolower(names(of))
  num_basins <- get_basin_numbers(runoff_data)
  plot_titles <- paste(viscos_options("plot_title"),num_basins,sep = "")
  user_of <- of[[tolower(of_name)]] 
```

Afterwards a distinction is made if the the overall 
objective functions  or period-wise calculated objective functions have been
extracted. The check is done be checking if the first dimenion of `user_of` is
NULL (*overall of*) or has a dimension (*period-wise of*). Then `user_of` is
changed in two steps so that it can be plotted easily with
[ggplot2](http://ggplot2.org/). Step (I) transforms `user_of` into a data.frame.
To work properly the *overall of* vector has to be transposed before it is 
transfomred. The *period-wise of* can be transformed directly.  Before the 
transformation into a data.frame `pmax(.,viscos_options("limits")[1])` is 
applied to set the lowest values to the one specified in lower limits.
Step (II) then adds the information about the periods to `user_of`. 
```{r}
  only_one_period_or_total <- is.null(dim(user_of)[1])
  # (I)
  if ( only_one_period_or_total ) {
    user_of %<>% 
      t() %>%
      pmax(.,viscos_options("limits")[1]) %>%
      as.data.frame()
    names(user_of) <- plot_titles
  } else {
    user_of %<>%
      pmax(.,viscos_options("limits")[1]) %>%
      as.data.frame()
    names(user_of) <- plot_titles
  }
  # (II)
  if ( only_one_period_or_total ) {
    user_of$periods <- 1
  } else {
    user_of$periods <- unique(runoff_data$period) %>% 
      extract(. > 0)
  }
```

Once `user_of` is suitably transformed the defacto plot function , 
named `plot_function` is defined. This function will then be used in a loop 
to plot the individual plot. 
```{r}
  # define plot function: 
  plot_function <- function(k) {
    ggplot(data = user_of,environmnet = environment()) +
    geom_bar(stat = "identity",
             position = "identity",
             aes_string(x = "periods", y = plot_titles[k],fill = plot_titles[k])) +
    theme_light(base_size = 15) +
    ggtitle(plot_titles[k]) +
    xlab(viscos_options("xlab")) +
    ylab( gsub("_.*","", of_name) ) + # remove everything after "_" 
    scale_y_continuous(limits = viscos_options("limits")) +
    theme(legend.position = "none",
          axis.text.x = element_text(angle = 50, hjust = 1)
          ) + 
    scale_fill_gradient2(space = "Lab",
                         low = viscos_options("color_of_low"),
                         mid = viscos_options("color_of_mid"),
                         high = viscos_options("color_of_high"),
                         midpoint = viscos_options("midpoint"),
                         limits = viscos_options("limits") )
  }
```

```{r}
  # apply plot function over all columns in user_of
  eval_size <- ncol(user_of)
  list_of_barplots <- lapply(1:eval_size,
                    function(k) plot_function(k)
                    )
  return(list_of_barplots)
}

```

