---
title: "Code for listplot functions"
author: 
- "Daniel Klotz, Johannes Wesemann, Mathew Herrnegger"
date: "14 Juli 2016"
output: html_document
---

```{r setup, include=FALSE, purl=FALSE}
  knitr::opts_chunk$set(eval = FALSE, tidy = FALSE)
```

# Code 
In this section the code of the plotlist functions are defined

## The listplot function family
The wrapper function for the different plotlists of the objective functions.
The code is: 

```{r}
#' plotlists for the objective functions 
#'
#' To be written soon
#' @export
```

```{r}
listplot <- function(runoff_data,
                     kind = "barplot") {
  if (kind == "barplot") {
    plot_list <- serve.plotlist_barplot(runoff_data)
  } else {
    stop("no such option for the kind of plot")
  }
  return(plot_list)
}
  

```


Users can define the to-be plotted objective function with the character 
`of_name`. 

The first step is the calculation of the objective functions (`of`) via 
[`extract_objective_functions`](LP-computing_of.html).
Then some housekeeping needs to be undertaken. For the computations and 
selections all names are set to lower case (`tolower`), because **visCOS** works
case-insensitive. The basin enumeration `num_basins` is obtained with
the function [`get_basin_numbers`](LP-helpers.html). The plot-titles 
of the individual barplots are build by gathering the main title from the 
[global option](LP-viscos_options.html) *"plot_title"* and concatenating it with 
`num_basins`. The user defined objective function (`user_of`) is extracted from 
`of` by using the user defined selection `of_name`. Again, in order to avoid 
lower-upper case distinctions the `of_name` string is allways `set tolower`. 

```{r}
serve.plotlist_barplot <- function(runoff_data) {
  require("ggplot2", quietly = TRUE)
  require("magrittr", quietly = TRUE)
  #################
  # calculate objective functions
  of <- extract_objective_functions(runoff_data)
  # 
  num_basins <- ncol(of) %>% subtract(1)
  # define plot function: 
  melted_user_of <- suppressMessages( reshape2::melt(of) )%>% 
    mutate(of_group = ifelse(of %>% as.character %>% startsWith(.,"NSE"),"NSE",
                             ifelse(of %>% as.character %>% startsWith(.,"KGE"),"KGE",
                                    ifelse(of %>% as.character %>% startsWith(.,"CORR"),"CORR",
                                           ifelse(of %>% as.character %>% startsWith(.,"pBIAS"),"pBIAS",NA))))) 
  groupings_of <- melted_user_of$of_group %>% unique
```

```{r}
  # plot list function
  plotlist_fun <- function(grouping) {
    of_to_plot <- melted_user_of %>% filter( of_group == grouping)
    if (grouping == "pBIAS") {
    plot_list <- ggplot(data = of_to_plot) +
      geom_bar(stat = "identity",
               position = "identity",
               aes(x = of,
                   y = value)) +
      ggtitle(grouping) +
      facet_wrap(~ variable, ncol = 1) + 
      ylim(c(-100,100))
    return(plot_list)
    } else {
    plot_list <- ggplot(data = of_to_plot) +
      geom_bar(stat = "identity",
               position = "identity",
               aes(x = of,
                   y = value)) +
      facet_wrap(~ variable, ncol = 1) + 
      ggtitle(grouping) +
      ylim(viscos_options("limits"))
    return(plot_list)
    }
  }
```

```{r}
  # map plotlist_fun
  plot_list <- lapply(groupings_of,plotlist_fun)
  names(plot_list) <- c("NSE","KGE","pBIAS","CORR")
  return(plot_list)
}
```


