---
title: "Make Time Aggregates"
author: 
- "Daniel Klotz, Johannes Wesemann"
date: "22 September 2016"
output:
  html_document:
    number_sections: yes
---



```{r setup, include=FALSE, purl=FALSE}
  knitr::opts_chunk$set(eval = FALSE, tidy = FALSE, fig.align = 'center')
```

# Introduction
In hydrology it is often usefull to summarise the runoff with respect to a 
given time dimension. In **visCOS** this can be done by using the 
`aggregate_time` function. Thevfunction takes 
[COSERO data.frame](cook_data.html) and aggregates them according to a chosen
time dimension. Note, that the name of the dimension can be specified via the 
[options](LP-viscos_options.html).


```{r, eval = TRUE, purl=FALSE}
  require(ggplot2, quietly = TRUE)
  require(visCOS, quietly = TRUE)
```

```{r, eval = TRUE, purl=FALSE}
  runoff_data <- visCOS::get_runoff_example()
  runoff_aggregate_dd <- aggregate_time(runoff_data, "dd")
  ggplot(runoff_aggregate_dd) + 
      geom_line(aes(x = idx, y = value, col = obs_sim)) + 
      scale_colour_manual(values = c(viscos_options("color_data1"),
                                     viscos_options("color_data2"))) + 
      facet_wrap( ~ basin,ncol = 1) +
      theme_minimal()
```

```{r, eval = TRUE, purl=FALSE}
  runoff_aggregate_mm <- aggregate_time(runoff_data, "mm")
  ggplot(runoff_aggregate_mm) + 
      geom_line(aes(x = idx, y = value, col = obs_sim)) + 
      scale_colour_manual(values = c(viscos_options("color_data1"),
                                     viscos_options("color_data2"))) + 
      scale_x_discrete(limits = runoff_aggregate_mm$time_aggregate) +
      facet_wrap( ~ basin, scales= "free") + 
      theme_minimal()
```

```{r, eval = TRUE, purl=FALSE}
  runoff_aggregate_yyyy <- aggregate_time(runoff_data, "yyyy")
  ggplot(runoff_aggregate_yyyy) + 
    geom_point(aes(x = idx, y = value, col = obs_sim)) + 
    scale_colour_manual(values = c(viscos_options("color_data1"),
                                   viscos_options("color_data2"))) + 
    facet_wrap( ~ basin) +
    scale_x_discrete(limits = runoff_aggregate_yyyy$time_aggregate,  
                    labels = abbreviate) + 
    theme_minimal()
```

```{r, eval = TRUE, purl=FALSE}
  runoff_aggregate_yyyymm <- aggregate_time(runoff_data, "yyyy-mm")
  ggplot(runoff_aggregate_yyyymm) + 
    geom_line(aes(x = idx, y = value, col = obs_sim)) + 
    scale_colour_manual(values = c(viscos_options("color_data1"),
                                   viscos_options("color_data2"))) + 
    scale_x_discrete(limits = runoff_aggregate_yyyymm$time_aggregate,  
                     labels = abbreviate) +
    facet_wrap( ~ basin,ncol = 1) +
    theme_minimal() + 
    theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

# Code

```{r}
  #' Aggreage and plot
  #' 
  #' Hopefully a good description will follow 
  #'
  #' @import magrittr
  #' @import ggplot2
  #' @import pasta
  
  #' @export
  aggregate_time <- function( runoff_data, aggregation = "mm") {
```
The time aggregation is done by cutting the needed information out of the 
date-string. This is a bit ruff, but works nicely and seems to be more 
commonly used than you would expect. 
```{r}
    #
    cutting_bounds <- c(Inf,-Inf)
    if (grepl("dd",aggregation)) {
      cutting_bounds[1] <- min(9,cutting_bounds[1])
      cutting_bounds[2] <- max(11,cutting_bounds[2])
    }
    if (grepl("mm",aggregation)) {
      cutting_bounds[1] <- min(6,cutting_bounds[1])
      cutting_bounds[2] <- max(7,cutting_bounds[2])
    }
    if (grepl("yyyy",aggregation)) {
      cutting_bounds[1] <- min(1,cutting_bounds[1])
      cutting_bounds[2] <- max(4,cutting_bounds[2])
    }
```

```{r}
    ######  define helpers
    regex_for_runoff_selection <- viscos_options("name_data1") %|%  viscos_options("name_data2")
    # aggregation function:
    aggregator_fun <- function(k,data_frame){
      the_aggregation <- aggregate(data_frame[[k]] ~ data_frame$date_selection, FUN = mean)
      return(the_aggregation[ ,2])
    }
    ##### 
    runoff_data %>% 
      visCOS::prepare_complete_date() %>% 
      visCOS::remove_chunk() -> full_runoff_data 
    runoff_with_aggreggation <- cbind.data.frame(
      full_runoff_data,
      date_selection = substr(full_runoff_data$posixdate,
                              cutting_bounds[1],
                              cutting_bounds[2]) %>% as.factor()
      )
    names_runoff_selection <- grep(
      regex_for_runoff_selection,
      names(runoff_with_aggreggation) %>% tolower, 
      value = TRUE
      )
    selected_runoff_rows <- grep(regex_for_runoff_selection,
                             names(runoff_with_aggreggation), 
                             ignore.case = TRUE)
    selected_runoff_rows %>% 
      sapply(.,function(x) aggregator_fun(x,runoff_with_aggreggation)) %>% 
      data.frame(idx = 1:nrow(.), 
                 time_aggregate = unique(runoff_with_aggreggation$date_selection),
                 .) %>% 
      set_names(., c("idx","time_aggregate",names_runoff_selection)) -> time_aggregate
    # 
    melted_time_aggregate <- time_aggregate %>% 
      reshape2::melt(., id.vars = c("idx","time_aggregate")) %>% 
      cbind.data.frame(., 
                       basin =  .$variable %>%
                         gsub(regex_for_runoff_selection,"",.) %>% 
                         gsub("\\D","",.) %>%
                         as.integer, 
                       obs_sim = .$variable %>% 
                         gsub(viscos_options("name_data1") %&% ".*",viscos_options("name_data1"),.) %>% 
                         gsub(viscos_options("name_data2") %&% ".*",viscos_options("name_data2"),.))
    return(melted_time_aggregate)
  }
```



