---
title: "Code for plotlist of"
author: 
- "Daniel Klotz, Johannes Wesemann, Mathew Herrnegger"
date: "14 Juli 2016"
output: html_document
---

```{r setup, include=FALSE, purl=FALSE}
  knitr::opts_chunk$set(eval = FALSE, tidy = FALSE)
```



The description for the documentation is: 
```{r}
#' hydyearly list of expaned barpltos 
#'
#' creates a plotlist of bar-plots for a the provided objective functions (of)
#' @export
```

The function header is: 
```{r}
serve.plotlist_periodOF <- function(runoff_data, of_name = "NSE_period") {
  require("ggplot2", quietly = TRUE)
  require("magrittr", quietly = TRUE)
  #################
```

The first step is the calculation of the objective functions (`of`) via 
[`extract_objective_functions`](LP-computing_of.html).
Then some housekeeping needs to be undertaken. For the computations and 
selections all names are set to lower case (`tolower`), because **visCOS** works
case-insensitive. The basin enumeration `num_basins` is obtained with
the function [`get_basin_numbers`](LP-helpers.html). The plot-titles 
of the individual barplots are build by gathering the main title from the 
[global option](LP-viscos_options.html) *"plot_title"* and concatenating it with 
`num_basins`. 
```{r}
  # calculate objective functions
  of <- extract_objective_functions(runoff_data)
  # 
  names(of) <- tolower(names(of))
  num_basins <- get_basin_numbers(runoff_data)
  plot_titles <- paste(viscos_options("plot_title"),num_basins,sep = "")
  temp_of <- of[[tolower(of_name)]] 
```

```{r}
  only_one_period_or_total <- is.null(dim(temp_of)[1])
  if ( only_one_period_or_total ) {
    ofun_period <- temp_of %>% 
      t() %>%
      pmax(.,viscos_options("lb_cut")) %>%
      as.data.frame()
    names(ofun_period) <- plot_titles
  } else {
    ofun_period <- temp_of %>%
      pmax(.,viscos_options("lb_cut")) %>%
      as.data.frame()
    names(ofun_period) <- plot_titles
  }
  #
  eval_size <- ncol(ofun_period)
  if ( only_one_period_or_total ) {
    ofun_period$periods <- 1
  } else {
    ofun_period$periods <- unique(runoff_data$period) %>% extract(. > 0)
  }
  # define plot function: 

  plot_function <- function(k) {
    ggplot(data = ofun_period,environmnet = environment()) +
    geom_bar(stat = "identity",
             position = "identity",
             aes_string(x = "periods", y = plot_titles[k],fill = plot_titles[k])) +
    theme_light(base_size = 15) +
    ggtitle(plot_titles[k]) +
    xlab(viscos_options("xlab")) +
    ylab( gsub("_.*","", of_name) ) + # remove everything after "_" 
    scale_y_continuous(limits = viscos_options("limits")) +
    theme(legend.position = "none",
          axis.text.x = element_text(angle = 50, hjust = 1)
          # clockwise from above ,plot.margin = grid::unit(c(0.2,0.5,0.2,0.5), "cm")
          ) + 
    scale_fill_gradient2(space = "Lab",
                         low = viscos_options("color_of_low"),
                         mid = viscos_options("color_of_mid"),
                         high = viscos_options("color_of_high"),
                         midpoint = viscos_options("midpoint"),
                         limits = viscos_options("limits") )
  }
  # apply plot function over all available stuff
  list_of_barplots <- lapply(1:eval_size,
                    function(k) plot_function(k)
                    )
  return(list_of_barplots)
}

```

