---
title: "visCOS: Example 1"
output: 
  html_document:
    highlight: pygments
---
```{r, echo = FALSE, message=FALSE}
  require("visCOS")
  require("magrittr")
```

## Intro to the pour_ functions
As a first step let us get some exemplary runoff and take a look at it. Within **visCOS** can be done via the `pour_runoff_example` command. 
```{r}
  head( pour_runoff_example() ) # get example runoff
```

The first five rows of the example are used to encapsulate the date in the **COSdate** format (see: xxx). We can also not that there are ome unnessary colums named QOSI. We can easily remove the not needed columns with `prepare_removeChunk`, which will remove all rows except those, which names begin with **qobs**, **qsim**, are in the **COSdate** format, or are named **posixdate** and **period** (the name testing is not case-sensitive) xxx

xxx Explain that one can test and play with the runoff



xxx a bit more tricky:

```{r}
  spinup_a <- pour_spinup(filepath = "dummy",pattern = "a")
  spinup_b <- pour_spinup("dummy","b")
  spinup_c <- pour_spinup("dummy","c")
  # note: there is no "d"-pattern in the dummy variable
    c(spinup_a,spinup_b,spinup_c)
  
```


xxx
```{r}
  pour_plt_ctrl()
```




